---
title: "Conservation Status"
author: "Florian Mayer, DBCA"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ckanr)
library(readr)
library(magrittr)
library(knitr)
library(dplyr)
library(tidyr)
library(purrr)
source("helpers.R")
if (file.exists("~/.Rprofile")) source("~/.Rprofile")
ckanr::ckanr_setup(url = Sys.getenv("CKAN_URL"), key = Sys.getenv("CKAN_API_KEY"))
local_api <- "http://localhost:8220/api/1/"
```

# Context
This workbook explores assumptions around conservation status management using TSC data.
Assumptions are formulated as quoted statements, and their validity is explored through data analysis.

# Glossary
* Scope: WA=State, FED=Commonwealth/Federal, INT=International.
* [IUCN](https://www.iucn.org/): International Union for Conservation of Nature. 
  IUCN publish the [Red List](http://www.iucnredlist.org/) which has an 
  [API](http://apiv3.iucnredlist.org/).
* "Gazettal" the assignment of a conservation status to a taxnonomic name.

# Data
Nomenclature: 

* `tfa` = Threatened and priority fauna
* `tfl` = Threatened and priority flora
* `tec` = Threatened and priority communities
* `gaz` = conservation status gazettal
* `cl` = conservation list
* `cat` = conservation category (EN, VU etc)
* `crit` = IUCN criteria for conservation status gazettal (2biii)
* `wa` = state
* `cmw` = commonwealth
* `int` = international

```{r load_data, , eval=T, message=FALSE}
load_ckan_csv <- .  %>% resource_show() %>% extract2("url") %>% read_csv()
tfa_cl <- "2793b1a8-eb55-41e7-8f8f-9cbc79a01243" %>% load_ckan_csv
tfa_gaz <- "ac089664-de5a-4efa-af3c-ada0ec77ef1f" %>% load_ckan_csv
tfa_cc <- "2793b1a8-eb55-41e7-8f8f-9cbc79a01243" %>% load_ckan_csv

tfl_gaz_list <- "0fbd9080-21d7-4fb1-ae43-4cc5791862c3" %>% load_ckan_csv
tfl_gaz_prio <- "668d2bd2-02bf-43fc-ad88-439185013d15" %>% load_ckan_csv

tec_cl <- NULL  # list1: home-brewed, list2: IUCN RLE
tec_cat <- NULL
tec_crit <- NULL
tec_gaz_cat <- "f0085651-8c38-496e-a6bd-1adc7ae2f06a" %>% load_ckan_csv
tec_gaz_crit <- "9483f333-97fd-4975-a799-d44bf8bd252a" %>% load_ckan_csv
```

# International conservation listing
"The Department does not interact with international conservation listings."

Discuss: 

* DBCA does not actively inform IUCN of WA gazettal.
* Should DBCA publish WA conservation status to IUCN and the general public?
* DBCA only needs a read-only copy of IUCN status.
* DBCA currently manually maintains a copy of IUCN conservation status.
* There is no reason not to automate import of IUCN statuses as long as the taxonomic names are correct.

# Conservation lists
## Fauna
Conservation lists and categories available:
```{r tfa_cl, eval=T}
tfa_cl %>% select(list_code, status_code, status) %>% kable(.)
```

## Flora
Conservation lists used:
```{r tfl_cl, eval=T}
unique(tfl_gaz_list$list_code)
```

## TEC
TEC has only the DBCA-internal conservation list, which will be superseded by the IUCN RLE.

```{r tec_cl}
tec_gaz_cat %>% select(category_type_code, category_type) %>% unique() %>% mutate(listcode="DBCA_RLE")
```

Wendy and Val advise:

* A proponent (TEC group or external expert) nominates a com as prio or threatened.
* The TECSC (TEC sci committee) approves the com nomination.
* The Minister endorses the nomination.

* The conservation categories 
  [we have](https://data.dpaw.wa.gov.au/dataset/threatened-ecological-communities-database/resource/aa2cc831-3c4d-408f-bb5b-61457b48e304) are home-cooked.
* By September, we are encouraged to use IUCN statuses.
* We need to preserve the history of statuses.
* IUCN categories and criteria meaning does not match DBCA criteria meaning.
* IUCN RLE (red list for ecosystems) defines categories (EN etc) and criteria (2Biii)
* DBCA categories were defined in absence of any other official categories in the mid 1990s,
  DBCA categories and criteria were signed off by the minister in the early 2000s for 69 com.
* With BCA going into effect ca Sep 2018, 
  new TEC nomionations will use IUCN RLE categories and criteria. 
  Existing cat (and possibly crit) will be rolled over into BCA and replaced by RLE over time.

Ask Val:
* When do we notify FED of TEC gazettal?
* When do FED notify DBCA of their gazettal?


## Combined

Conservation categories used in legacy data:

```{r wa_cat_used, eval=T}
cs_used <- list(
    tfa = unique(tfa_gaz$status_code),
    tfl = unique(tfl_gaz_list$status_code),
    tec = unique(tec_gaz_cat$category_type_code)
)
cs_used

# tfa_and_tec <- intersect(cs_used$tfa, cs_used$tec)
# tfa_not_tec <- setdiff(cs_used$tfa, cs_used$tec)

```

# Gazettals
A Conservation Status Gazettal is the assignment of one or several conservation criteria
from one or several conservation lists to a taxon, optionally justified by a number of IUCN criteria.

## Fauna

```{r tfa_gaz, eval=T}
tfa_gaz %>% head(.) %>% kable(.)
```

## Flora

```{r tfl_gaz, eval=T}
tfl_gaz_list %>% head(.) %>% kable(.)
```

## TEC

```{r tec_gaz, eval=T}
tec_gaz_cat %>% head(.) %>% kable(.)
```

# Taxonomic changes and Gazettal
Is the following correct?

* If a name splits/merges, cons status applies to successors interim until successors are gazetted or delisted.
* On the rollover date from WAWCA to BCA, BCA status codes replace WAWCA codes without date overlap.
* There is an exact 1:1 mapping of WAWCA to BCA status codes.
* IU will be split into four migratory statuses, which are not mutually exclusive.
* ActionPlan gazettals are read by an officer (offline and outside TSC), who then may choose 
  to initiate a WAGazettal.
* There is no need for separate ActionPlan gazettals to be in TSC, as they may or may not result 
  in regular WAGazettals.

* Federal gazettal is started from our end after TSSC approval of a WAGazettal.
* If Federal gazettal is successful, we are notified.
* TSC does not need to keep track of intermediary federal gazettal approval stages.

* International gazettal is read-only to TSC.

```{r wa_gaz_combined, eval=T}
tfa_wawca <- tfa_cl %>% filter(list_code %in% c("WAWCA", "SPFN", "WAPF"))

filter_tfa <- function(x) filter(tfa_cl, list_code == x) %>% select(list_code, status_code)
tfa_epbc <- "EPBC" %>% filter_tfa %>% rename(EPBC = list_code)
tfa_iucn2012 <- "IUCN(2012)" %>% filter_tfa %>% rename(IUCN2012 = list_code)
tfa_iucn2001 <- "IUCN(2001)" %>% filter_tfa %>% rename(IUCN2001 = list_code)
tfa_iucn1994 <- "IUCN(1994)" %>% filter_tfa %>% rename(IUCN1994 = list_code)
tfa_redlist <- "RedList" %>% filter_tfa %>% rename(RedList = list_code)
tfa_ap <- "ActionPlan" %>% filter_tfa %>% rename(ActionPlan = list_code)

tfa_status_comp <- tfa_wawca %>%
    select(list_code, status_code) %>%
    full_join(tfa_epbc, by="status_code") %>%
    full_join(tfa_iucn2012, by="status_code") %>%
    full_join(tfa_iucn2001, by="status_code") %>%
    full_join(tfa_iucn1994, by="status_code") %>%
    full_join(tfa_redlist, by="status_code") %>%
    full_join(tfa_ap, by="status_code") 

tfa_status_comp %>% kable(.)
```


# Recovery Plans
* Each plan has a team of users with rec plan specific roles and extra information. 
  We need to maintain this Membership info.
* TSC does not need to maintain teams - we can create rec plan team memberships for each team member.
* Rec plans can pertain to one or several taxa.

# Data upload
## Conservation Lists, Categories, Criteria
Fauna cons lists and categories are uploaded to TSC.

```{r data_upload_tsc, eval=F}
# Fauna Conservation lists alone
# "9c3f8cac-e8ed-4713-add2-067b83f410df" %>% 
#     load_ckan_csv %>% 
#     dplyr::transmute(
#         code = list_code,
#         label = list_name
#         ) %>% 
#     wastdr::wastd_POST("conservationlist", api_url = local_api)

# Fauna conservation lists and categories
"2793b1a8-eb55-41e7-8f8f-9cbc79a01243" %>% 
    load_ckan_csv %>%                       # Same as tfa_cl
    dplyr::transmute(                       # Cons category col names
            code = status_code, 
            label = status, 
            conservation_list = list_code) %>% 
    dplyr::group_by(conservation_list) %>%  # One line per cons list
    tidyr::nest() %>%                       # One list col with all cons categories
    dplyr::rename(                          # Cons list col names as per TSC API
        code = conservation_list,
        conservationcategory_set = data
    ) %>% 
    transpose(.) %>%                        # Transpose to let iteration return rows of code and conscat_set
    wastdr::wastd_POST(., serializer = "conservationlist")  # , api_url = local_api)

tec_gaz_cat %>% 
    select(category_type_code, category_type) %>% 
    unique() %>% 
    mutate(list_code="DBCA_RLE") %>%
    dplyr::transmute(                       # Cons category col names
            code = category_type_code, 
            label = category_type, 
            conservation_list = list_code) %>% 
    dplyr::group_by(conservation_list) %>%  # One line per cons list
    tidyr::nest() %>%                       # One list col with all cons categories
    dplyr::rename(                          # Cons list col names as per TSC API
        code = conservation_list,
        conservationcategory_set = data
    ) %>% 
    transpose(.) %>%                        # Transpose to let iteration return rows of code and conscat_set
    wastdr::wastd_POST(., serializer = "conservationlist")  # , api_url = local_api)
```

This workbook is uploaded to the data catalogue.

```{r data_upload_ckan}
d <- ckanr::package_show("wacensus")
# r <- resource_create(package_id = d$id, name = "Conservation Status Analysis", 
# format = "HTML", upload = "conservation_status.html")
ckanr::resource_update("3d875d8e-8192-4312-a749-9dd04dd6799b", "conservation_status.html")
```
