---
title: "WACensus ETL"
author: "Florian Mayer, DBCA"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
  pdf_document: default
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(magrittr)
library(tidyverse)
library(skimr)
library(leaflet)
library(Hmisc)
library(ckanr)
library(wastdr)
source("helpers.R")
wastdr::wastdr_setup()
ckanr::ckanr_setup(url = Sys.getenv("CKAN_URL"), key = Sys.getenv("CKAN_API_KEY"))

# devtools::install_github("eblondel/geosapi")
# library(geosapi)

local_api <- "http://localhost:8220/api/1/"
```

# Context

* WACensus, an Oracle database, is the point of truth for WA texonomic names at DBCA.
* KMI publishes a few WACensus views as GeoServer views.
* BioSysTT maintains a copy of those WACensus views.

This workbook serves to update BioSysTT via its API from KMI's WACensus views.


# Helpers
Some helpers are defined.

```{r read_data_helpers}
kmi_getFeature <- function(layer_name = "dpaw:herbie_hbvnames",
                           url = "https://kmi.dbca.wa.gov.au/geoserver/dpaw/ows") {
  ua <- httr::user_agent("http://github.com/dbca-scarab-scripts")
  url <- "https://kmi.dbca.wa.gov.au/geoserver/public/ows"
  query <- list(
    service = "WFS",
    version = "2.0.0",
    request = "GetFeature",
    typeName = layer_name,
    outputFormat = "application/json"
  )
  url %>% httr::GET(ua, query = query) %>% httr::content(.)
}

wastd_POST <- function(data,
                       serializer = "names",
                       api_url = wastdr::get_wastdr_api_url(),
                       api_token = wastdr::get_wastdr_api_token(),
                       api_un = wastdr::get_wastdr_api_un(),
                       api_pw = wastdr::get_wastdr_api_pw(),
                       verbose=FALSE) {
  ua <- httr::user_agent("http://github.com/parksandwildlife/turtle-scripts")
  if (!is.null(api_token)) {
    auth <- httr::add_headers(c(Authorization = api_token))
  } else {
    auth <- httr::authenticate(api_un, api_pw, type = "basic")
  }
  url <- paste0(api_url, serializer, "/")
  if (verbose==TRUE) message("[wastd_POST] url ", url)
  res <- httr::POST(url, auth, ua, encode = "json", body = data)
  if (verbose==TRUE) message("[wastd_POST] status ", res$status)
  structure(
    list(
      serializer = serializer,
      response = res
    ),
    class = "wastd_api_post_response"
  )
}

print.wastd_api_post_response <- function(x, ...) {
  cat("<WAStD API POST response ", x$serializer, ">\n",
    "URL ", x$response$url, "\n",
    "Date ", x$response$headers$date, "\n",
    "Status ", x$response$status_code, "\n",
    sep = ""
  )
  invisible(x)
}

wastd_upsert_geojson <- function(kmi_featurecollection,
                                     serializer = "names",
                                     api_url = wastdr::get_wastdr_api_url(),
                                     chunksize = 500,
                                     verbose = TRUE) {
  message("[wastd_upsert_geojson] Updating ", api_url, serializer, "...")
  props <- purrr::map(kmi_featurecollection[["features"]], "properties")
  # purrr::map(props, wastd_POST, api_url = api_url) # One by one - very slow. Faster:
  len <- length(props)
  for (i in 0:(len / chunksize)) {
    start <- (i * chunksize) + 1
    end <- min((start + chunksize) - 1, len)
    message("[wastd_upsert_geojson] Processing feature ", start, " to ", end)
    props[start:end] %>%
      purrr::map(., purrr::flatten) %>%
      wastd_POST(., serializer = serializer, api_url = api_url, verbose = verbose)
  }
  message("[wastd_upsert_geojson] Finished. ", len, " records updated.")
}
```

## Read all data from KMI

174 MB load in 5 seconds.

```{r read_data}
hbvnames <- kmi_getFeature(layer_name = "public:herbie_hbvnames_public")
hbvsupra <- kmi_getFeature(layer_name = "public:herbie_hbvsupra_public")

# kmi <- GSManager$new(
#   url = "https://kmi.dbca.wa.gov.au/geoserver/public/ows",
#   user = Sys.getenv("AD_UN"),
#   pwd = Sys.getenv("AD_PW"),
#   logger = "INFO"
# )
#
# hbvnames <- kmi$getLayer("herbie_hbvnames_public")
```

# Write data to BioSys TT

## Example: update one Taxon

```{r write_data_example}
namesprops <- purrr::map(hbvnames[["features"]], "properties")

wastd_POST(namesprops[[1]], serializer = "names", api_url = local_api)
wastd_POST(namesprops[[1]], serializer = "names")

supraprops <- purrr::map(hbvsupra[["features"]], "properties")
wastd_POST(supraprops[[1]], serializer = "supra", api_url = local_api)

```

## Batch upsert: Create or update all Taxa

Runs for about 50 minutes (one by one) or 25 minutes (after the upgrade to POST chunked).
names: 11:53 - 
roughly 1000 features per minute

```{r rock_and_roll, eval=FALSE}
wastd_upsert_geojson(hbvnames, serializer = "names", api_url = local_api, verbose = T)
wastd_upsert_geojson(hbvsupra, serializer = "supra", api_url = local_api, verbose = T)
```

# Upload to CKAN

```{r upload_ckan}
d <- ckanr::package_show("wacensus")
# r <- resource_create(package_id = d$id, name = "WACensus ETL to BioSysTT", format = "HTML", upload = "data_etl_wacensus.html")
ckanr::resource_update("2893ad10-dfbf-460d-a508-5195edfd9a89", "data_etl_wacensus.html")
```
