---
title: "WACensus ETL"
author: "Florian Mayer, DBCA"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(magrittr)
library(tidyverse)
library(skimr)
library(leaflet)
library(Hmisc)
library(ckanr)
library(wastdr)
source("helpers.R")
if (file.exists("~/.Rprofile")) source("~/.Rprofile")
ckanr::ckanr_setup(url = Sys.getenv("CKAN_URL"), key = Sys.getenv("CKAN_API_KEY"))
local_api <- "http://localhost:8220/api/1/"
```

# Context

* WACensus, an Oracle database, is the point of truth for WA texonomic names at DBCA.
* KMI publishes a few WACensus views as GeoServer views.
* TSC maintains a copy of those WACensus views.

This workbook serves to update TSC via its API from KMI's WACensus views.

# Write data to BioSys TT

## Example: update one Taxon

```{r write_data_example, eval=FALSE}
hbvnames <- kmi_getFeature(layer_name = "public:herbie_hbvnames_public")
namesprops <- purrr::map(hbvnames[["features"]], "properties")
wastd_POST(namesprops[[1]], serializer = "names", api_url = local_api)
```

## Batch upsert: Create or update all Taxa
Local TSC instance (dev):

```{r rock_and_roll_local, eval=FALSE}
"public:herbie_hbvnames_public" %>% 
    gs_getFeature %>% 
    upsert_geojson(serializer = "names", verbose = T, api_url = local_api)
"public:herbie_hbvsupra_public" %>% 
    gs_getFeature %>% 
    upsert_geojson(serializer = "supra", verbose = T, api_url = local_api)
"public:herbie_hbvgroups_public" %>% 
    gs_getFeature %>% 
    upsert_geojson(serializer = "groups", verbose = T, api_url = local_api)
"public:herbie_hbvfamilies_public" %>% 
    gs_getFeature %>% 
    upsert_geojson(serializer = "families", verbose = T, api_url = local_api)
"public:herbie_hbvgenera_public" %>% 
    kmi_getFeature %>% 
    upsert_geojson(serializer = "genera", verbose = T, api_url = local_api)
"public:herbie_hbvspecies_public" %>% 
    gs_getFeature %>% 
    upsert_geojson(serializer = "species", verbose = T, api_url = local_api)
"public:herbie_hbvvernaculars_public" %>% 
    gs_getFeature %>% 
    upsert_geojson(serializer = "vernaculars", verbose = T, api_url = local_api)
"public:herbie_hbvxrefs_public" %>% 
    gs_getFeature %>% 
    upsert_geojson(serializer = "xrefs", verbose = T, api_url = local_api)
"public:herbie_hbtparents_public" %>% 
    gs_getFeature %>% 
    upsert_geojson(serializer = "parents", verbose = T, api_url = local_api)
```

Production TSC instance.

```{r rock_and_roll_production, eval=FALSE}
"public:herbie_hbvnames_public" %>% gs_getFeature %>% upsert_geojson(serializer = "names")
"public:herbie_hbvsupra_public" %>% gs_getFeature %>% upsert_geojson(serializer = "supra")
"public:herbie_hbvgroups_public" %>% gs_getFeature %>% upsert_geojson(serializer = "groups")
"public:herbie_hbvfamilies_public" %>% gs_getFeature %>% upsert_geojson(serializer = "families")
"public:herbie_hbvgenera_public" %>% gs_getFeature %>% upsert_geojson(serializer = "genera")
"public:herbie_hbvspecies_public" %>% gs_getFeature %>% upsert_geojson(serializer = "species")
"public:herbie_hbvvernaculars_public" %>% gs_getFeature %>% upsert_geojson(serializer = "vernaculars")
"public:herbie_hbvxrefs_public" %>% gs_getFeature %>% upsert_geojson(serializer = "xrefs")
"public:herbie_hbtparents_public" %>% gs_getFeature %>% upsert_geojson(serializer = "parents")
```

## Updating users
Given a CSV with columns `name`, `phone`, `email` and `role`, we can update users as follows.

```{r update_users, eval=F}
readr::read_csv("data/staff.csv") %>% wastd_POST(., serializer = "users", verbose = TRUE)
```

The username is generated from the `snake_case`'d field `name`.

# Upload to CKAN

```{r upload_ckan}
d <- ckanr::package_show("wacensus")
# r <- resource_create(package_id = d$id, name = "WACensus ETL to BioSysTT", format = "HTML", upload = "data_etl_wacensus.html")
ckanr::resource_update("2893ad10-dfbf-460d-a508-5195edfd9a89", "data_etl_wacensus.html")
```
