---
title: "WACensus ETL"
author: "Florian Mayer, DBCA"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
  pdf_document: default
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(magrittr)
library(tidyverse)
library(skimr)
library(leaflet)
library(Hmisc)
library(ckanr)
library(wastdr)
source("helpers.R")
wastdr::wastdr_setup()
ckanr::ckanr_setup(url = Sys.getenv("CKAN_URL"), key = Sys.getenv("CKAN_API_KEY"))

local_api <- "http://localhost:8220/api/1/"
```

# Context

* WACensus, an Oracle database, is the point of truth for WA texonomic names at DBCA.
* KMI publishes a few WACensus views as GeoServer views.
* BioSysTT maintains a copy of those WACensus views.

This workbook serves to update BioSysTT via its API from KMI's WACensus views.


# Helpers
Some helpers are defined.

```{r read_data_helpers}
kmi_getFeature <- function(
    layer_name = "dpaw:herbie_hbvnames", 
    url = "https://kmi.dbca.wa.gov.au/geoserver/dpaw/ows"){
    ua <- httr::user_agent("http://github.com/dbca-scarab-scripts")
    url <- "https://kmi.dbca.wa.gov.au/geoserver/dpaw/ows"
    query <- list(
        service = "WFS",
        version = "2.0.0",
        request = "GetFeature",
        typeName = layer_name,
        outputFormat="application/json"
    )
    res <- httr::GET(url, ua, query = query)
    httr::content(res)
}

wastd_POST <- function(
    data,
    serializer = "taxonomy",
    api_url = wastdr::get_wastdr_api_url(),
    api_token = wastdr::get_wastdr_api_token(),
    api_un = wastdr::get_wastdr_api_un(),
    api_pw = wastdr::get_wastdr_api_pw(),
    verbose=FALSE
){
    ua <- httr::user_agent("http://github.com/parksandwildlife/turtle-scripts")
    
    if (!is.null(api_token)) {
        auth <- httr::add_headers(c(Authorization = api_token))
      } else {
        auth <- httr::authenticate(api_un, api_pw, type = "basic")
      }
    
    url <- paste0(api_url, serializer, "/")
    
    res <- httr::POST(url, auth, ua, encode = "json", body = data)
    
    if (verbose) message("[wastd_POST] status ", res$status)
    
     structure(
        list(
          serializer = serializer,
          response = res
        ),
        class = "wastd_api_post_response"
    )
}

print.wastd_api_post_response <- function(x, ...) {
  cat("<WAStD API POST response ", x$serializer, ">\n",
    "URL ", x$response$url, "\n",  
    "Date ", x$response$headers$date, "\n",
    "Status ", x$response$status_code, "\n",
    sep = ""
  )
  invisible(x)
}

wastd_batch_update_taxon <- function(
    kmi_featurecollection,
    api_url = wastdr::get_wastdr_api_url(),
    verbose = T
    ){
    message("[wastd_batch_update_taxon] Updating Taxon...")

    props <- purrr::map(kmi_featurecollection[["features"]], "properties")
    
    # One by one - very slow.    
    # purrr::map(props, wastd_POST, api_url = api_url)
    
    chunksize <- 100
    len <- length(props)

    for (i in 0:(len/chunksize)){
        start <- (i * chunksize) + 1
        end <- min((start + chunksize) - 1, len)
        message("Processing feature ", start, " to ", end)
        
        props[start:end] %>% 
            purrr::map(., purrr::flatten) %>%
            wastd_POST(., api_url = api_url, verbose = verbose)
    }
    
    message("[wastd_batch_update_taxon] Finished. ", len, " taxa updated.")
}
```

## Read all data from KMI

174 MB load in 5 seconds.

```{r read_data}
f <- kmi_getFeature(layer_name = "dpaw:herbie_hbvnames")
```

# Write data to BioSys TT

## Example: update one Taxon

```{r write_data_example}
props <- purrr::map(f[["features"]], "properties")

# one_update <- wastd_POST(props[[1]], api_url = local_api, verbose=T)
one_update <- wastd_POST(props[[1]], verbose=T)

one_update$response %>% httr::content(.) %>% extract2("name")
one_update

# all_updates <- purrr::map(props, wastd_POST, api_url = local_api)
```

## Batch upsert: Create or update all Taxa

Runs for about 50 minutes (one by one) or 25 minutes (chunked).

```{r rock_and_roll, eval=FALSE}
updated <- wastd_batch_update_taxon(f)
```


# Upload to CKAN

```{r upload_ckan}
d <- ckanr::package_show("wacensus")
# r <- resource_create(package_id = d$id, name="WACensus ETL to BioSysTT", format="HTML", upload = "data_etl_wacensus.html")
ckanr::resource_update("2893ad10-dfbf-460d-a508-5195edfd9a89", "data_etl_wacensus.html")
```
