---
title: "Data ETL: TEC"
author: "Florian Mayer, DBCA"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (file.exists("~/.Rprofile")) source("~/.Rprofile")
source("helpers.R")
```

# Conventions
The following conventions shall be used to create table and column names in this ETL process.

* All names shall be snake_case, i.e. lower cased and underscore_separated.
* Primary key (PK) columns in lookup tables shall be prefixed by the table name. 
  This facilitates joining multiple lookups using unambiguous column names.
* Numeric PKs shall be named `_id`.
* Alphanumeric PKs (containing characters) shall be named `_code`.
* Abbreviations shall be avoided largely, with the exception of the ubiquitous 
  community (`com`) and occurrence (`occ`).

* All primary tables shall be joined to their lookup tables, and legacy lookup ID / code shall be discarded.
* Tables holding obsolete or duplicate data will be discarded.

# Context
[TEC dataset](https://data.dpaw.wa.gov.au/dataset/threatened-ecological-communities-database)
on the [departmental data catalogue](https://data.dpaw.wa.gov.au/).

The code for all SCB workbooks is under version control at [github](https://github.com/dbca-wa/scarab-scripts).


# Data
Data comes from four sources:

* An MS Access database dump from the original TEC Oracle database,
* three shapefiles (exported as GeoJSON) for TEC boundaries (exact site polygons),
  TEC buffers (boundaries buffered by a given amount), TEC sites (point locations
  inside boundaries referencing e.g. 10x10m quadrats).

```{r tec_data_load}
tec <- dl_mdbzip("598b1e70-5707-4dbc-a665-1189e1524ebe")
# tec_app <- dl_mdbzip("27d9b026-6933-4516-acf6-41edc6134adb")
tec_bdy <- geojsonsf::geojson_sf("data/tec_boundaries.geojson")
tec_buf <- geojsonsf::geojson_sf("data/tec_buffers.geojson")
tec_sit <- geojsonsf::geojson_sf("data/tec_sites.geojson")

# save(tec, tec_bdy, tec_buf, tec_sit, file=here::here("data", "tecdb.RData"))
# load(here::here("data", "tecdb.RData"))
```

# Configuration
Some tables store legacy configuration data. We won't need or import them.

* Settings: Datum preference
* Sp Master INI old: external db config
* mddDataTables/columns: external db tables
* mddMasterLists: external db config
* mddReportItems (empty)

# Taxonomy

* WA_* WACensus data copies
* jhb
* SPMASTER_OLD, Flora conservation codes
* Fauna taxa, Fauna conservation codes, Fauna Groups, Occ Fauna, Occ species, Species roles

Decision:

* Taxonomy is imported into TSC directly from WACensus, not from TEC.
* Taxonomy from TEC beyond NameID will be ignored.

# Spatial lookups
There are several groups of spatial entities which are maintained outside of the SCB ecosystem.
The link to WSCB entities can be inferred unambiguously through geospatial location and cached for performance.

* IBRA regions
* Beard: Districts within Regions within Provinces
* CALM: Occurrences within Districts within Regions. How are the occurrences inferred / added?
* CALM District Users
* DOLA locations
* Bush forever sites
* Zones (three broad zones in WA: rangelands, wheatbelt, southwest non-wheatbelt)
* Land use: Land Use, Land Use Codes, Occurrence Surr(?) Codes
* Tenures: Purposes, Tenure Types, Tenures, Occurrence Tenures, Occurrence Shires, Shires
* Maps: Maps, Occurrence Maps

# GIS exports
Some tables refer to externally managed spatial data:

* GIS_*
* Boundaries

# Spatial concepts
* Each occurrence lies within a `boundary` polygon.
* Not all occurrences have recorded boudaries.
* Boundaries (where existing) are buffered by a distance (default 500m) into a `buffer` polygon.
* Some buffers are reconstructed from a point plus a circular buffer, where boundaries are missing.
* There is one buffer for each occurrence.
* Point observations are somewhat misleadingly named `site`.
* Sites can serve:

  * To name a boundary
  * To create a buffer
  * To reference a point observation
  * To reference a terrestrial quadrat (default 10m x 10m) which is surveyed for species occurrences
* Some boundaries contain multiple sites.

There are 122970 sites, 120118 boundaries, and 121366 buffers.

# Communities
Abbreviation: `com`

```{r communities}
former_range <- tec$FORMER_RANGES %>%
  dplyr::transmute(
    former_range_code = FORMER.RANGE %>% as.character(),
    former_range = FORMER.RANGE.DESC %>% as.character()
  )
# make_coldef(former_range)

range_decline <- tec$RANGE_DECLINE %>%
  dplyr::transmute(
    range_decline_id = RANGE.DECLINE %>% as.integer(),
    range_decline = RANGE.DECLINE.DESC %>% as.character()
  )
# make_coldef(range_decline)

occ_decline <- tec$OCCURRENCES_DECLINE %>%
  dplyr::transmute(
    occ_decline_id = OCC.DECLINE %>% as.integer(),
    occ_decline = OCC.DECLINE.DESC %>% as.character()
  )
# make_coldef(occ_decline)

status <- tec$STATUS %>%
  dplyr::transmute(
    status_code = STATUS.CODE %>% as.character(),
    status = STATUS.DESC %>% as.character()
  )
# status %>% make_coldef()
# unique(status$status_code) # "BELIEVED"   "IDENTIFIED" - where used?

communities <- tec$COMMUNITIES %>%
  tibble::as.tibble() %>%
  dplyr::transmute(
    com_id = COM.NO %>% as.integer(),
    com_label = COM.ID %>% as.character(),
    com_name = COM.NAME %>% as.character(),
    former_range_code = COM.FORMER.RANGE %>% as.character(),
    range_decline_id = COM.RANGE.DECLINE %>% as.integer(),
    occ_decline_id = COM.OCC.DECLINE %>% as.integer(),
    com_tec_listing = COM.TEC.LISTING %>% as.integer(),
    com_original_area = COM.ORIG.AREA %>% as.double(),
    com_area_accuracy = COM.AREA.ACC %>% as.double(),
    preliminary = COM.PRELIMINARY %>% as.character(),
    status = COM.STATUS %>% as.character()
  ) %>%
  dplyr::left_join(former_range, by = "former_range_code") %>%
  dplyr::left_join(range_decline, by = "range_decline_id") %>%
  dplyr::left_join(occ_decline, by = "occ_decline_id") %>%
  dplyr::select(-former_range_code, -range_decline_id, -occ_decline_id)

communities %>% skim()
communities %>% DT::datatable(.)

com_short <- communities %>% dplyr::select(com_id, com_label, com_name)
```

## Community Recommendations

```{r com_recommendations}
recommendations <- tec$RECOMMENDATIONS %>%
  dplyr::transmute(
    recommendation_code = REC.CODE %>% as.character(),
    recommendation = REC.DESC %>% as.character()
  )

com_recommendations <- tec$COMMUNITY_RECOMMENDATIONS %>%
  dplyr::transmute(
    com_id = COM.NO %>% as.integer(),
    recommendation_id = CREC.NO %>% as.integer(),
    recommended_on = CREC.DATE %>% as.character(),
    recommendation_code = CREC.REC.CODE %>% as.character(),
    com_recommendation = CREC.DESC %>% as.character()
  ) %>%
  dplyr::left_join(com_short, by = "com_id") %>%
  dplyr::left_join(recommendations, by = "recommendation_code")

com_recommendations %>% DT::datatable(.)
```

## Community Publications

```{r com_publications}
publications <- tec$PUBLICATIONS %>%
  dplyr::transmute(
    publication_id = PUB.NO %>% as.integer(),
    title = PUB.TITLE %>% as.character(),
    author = PUB.AUTHOR %>% as.character()
    # empty:
    # published_on = PUB.DATE %>% parse_date_time(., orders = orders, tz = tz),
    # published_with = PUB.PLACE %>% as.character()
  )

com_publications <- tec$COMMUNITY_PUBLICATIONS %>%
  dplyr::transmute(
    com_id = COM.NO %>% as.integer(),
    publication_id = CP.PUB.NO %>% as.integer(),
  ) %>%
  dplyr::left_join(publications, by = "publication_id") %>%
  dplyr::left_join(com_short, by = "com_id")

com_publications %>% DT::datatable(.)
```

## Community Reviews

```{r com_reviews}
reviews <- tec$REVIEWS %>%
  dplyr::transmute(
    review_id = REV.NO %>% as.integer(),
    review_date = REV.DATE %>% parse_date_time(., orders = orders, tz = tz),
    review_panel = REV.PANEL %>% as.character()
  )

com_reviews <- tec$COMMUNITY_REVIEWS %>%
  dplyr::transmute(
    com_id = COM.NO %>% as.integer(),
    review_id = CR.REV.NO %>% as.integer()
  ) %>%
  dplyr::left_join(com_short, by = "com_id") %>%
  dplyr::left_join(reviews, by = "review_id")

com_reviews %>% DT::datatable(.)
```

## Community Actions

```{r com_actions}
actions <- tec$ACTIONS %>%
  dplyr::transmute(
    action_id = ACTION.CODE %>% as.character(),
    description = ACTION.DESC %>% as.character()
  )

com_actions <- tec$COMMUNITY_ACTIONS %>%
  dplyr::transmute(
    com_id = COM.NO %>% as.integer(),
    com_action_id = CACT.NO %>% as.integer(),
    com_action_date = CACT.DATE %>% parse_date_time(., orders = orders, tz = tz),
    action_id = CACT.ACTION.CODE %>% as.character(),
    com_action_description = CACT.DESC %>% as.character()
  ) %>%
  dplyr::left_join(com_short, by = "com_id") %>%
  dplyr::left_join(actions, by = "action_id")

com_actions %>% DT::datatable(.)
```

## Community History, Area Decline

```{r com_history}
area_decline <- tec$AREA_DECLINE %>%
  dplyr::transmute(
    area_decline_id = AD.CODE %>% as.integer(),
    area_decline = AD.DESC %>% as.character()
  )

com_history <- tec$HISTORY %>%
  dplyr::transmute(
    com_id = COM.NO %>% as.integer(),
    history_id = HIST.NO %>% as.character(),
    history_date = HIST.DATE %>% as.character(),
    area_decline_id = HIST.AD.CODE %>% as.integer(),
    comment = HIST.COMMENT %>% as.character()
  ) %>%
  dplyr::left_join(com_short, by = "com_id") %>%
  dplyr::left_join(area_decline, by = "area_decline_id")

com_history %>% DT::datatable(.)
```

# Occurrences

## Boundaries
Boundaries are named after the first site (see [site visits]) captured inside the respective boundary.

```{r tec_bdy , eval=F}
default_date <- lubridate::parse_date_time("1900-01-01 00:00:00", orders = "ymd HMS", tz = tz)
tec_bdy_cae <- tibble::tibble(
    community = tec_bdy$COM_ID,
    code=tec_bdy$FIRST_S_ID %||% tec_bdy$BDY_ID,
    name=tec_bdy$FIRST_S_ID %||% tec_bdy$BDY_ID,
    description = ifelse(
        is.na(tec_bdy$FIRST_S_ID),
        glue::glue("Boundary ID {tec_bdy$BDY_ID} named ",
                   "{tec_bdy$BDY_ID} for lack of sites within."),
        glue::glue("Boundary ID {tec_bdy$BDY_ID} named ",
                   "{tec_bdy$FIRST_S_ID} after first site observed within.")
    ),
    area_type = 10, # 10 bound, 11 buffer, 12 site point
    source = 13, # TEC bdy
    source_id = tec_bdy$OCC_UNIQUE,
    encountered_on=default_date,
    encountered_by=1,
    geom = tec_bdy$geometry %>% 
        sf::st_convex_hull(.) %>% 
        sf::st_buffer(., 0.00001) %>% 
        geojsonsf::sfc_geojson(.) 
    )
```

Some boundaries were digitised from a polygonised raster remote sensing product. This results in
a spray of numerous polygons, most of which were never visited in person.

Name and shame: 

```{r tec_cae_large}
numerous_communities <- tec_bdy_cae %>% 
  dplyr::group_by(community) %>% 
  dplyr::tally() %>% 
  dplyr::arrange(-n) %>% 
  filter(n>400) 

numerous_communities

fat_filter <- numerous_communities %>% 
  extract2("community") %>% 
  paste0(collapse = "|")
```

Import excluding the heavyweight communities with more than 400 features:
```{r tec_bdy_tsc, eval=F}
tec_bdy_cae %>% 
    filter(!grepl(fat_filter, community)) %>%
    chunk_post(serializer = "occ-community-areas", api_url = prod, verbose = T)
```

### Buffers

```{r tec_buf}
tec_buf_cae <- tibble::tibble(
    community = tec_buf$COM_ID,
    code=tec_buf$FIRST_S_ID %||% tec_buf$BDY_ID,
    name=tec_buf$FIRST_S_ID %||% tec_buf$BDY_ID,
    description = ifelse(
        is.na(tec_buf$FIRST_S_ID),
        glue::glue("Buffered Boundary ID {tec_buf$BDY_ID} named",
                   " {tec_buf$BDY_ID} for lack of sites within."),
        glue::glue("Buffered Boundary ID {tec_buf$BDY_ID} named",
                   " {tec_buf$FIRST_S_ID} after first site observed within.")
    ),
    area_type = 11, # 10 bound, 11 buffer, 12 site point
    source = 14, # TEC buf
    source_id = tec_buf$OCC_UNIQUE,
    encountered_on=default_date,
    encountered_by=1,
    geom = tec_buf$geometry %>% 
        sf::st_convex_hull(.) %>% 
        sf::st_buffer(., 0.00001) %>% 
        geojsonsf::sfc_geojson(.) 
    )
```

Import all but the numerous ones:

```{r tec_buf_tsc, eval=F}
tec_buf_cae %>% 
    dplyr::filter(!grepl(fat_filter, community)) %>%
    chunk_post(serializer = "occ-community-areas", api_url = prod, verbose = T)
```

### Sites

```{r tec_sit}
sit_com <- tec_sit %>% extract2("COM_ID") %>% unique
com_com <- com_short %>% extract2("com_label") %>% unique
site_name_typos <- setdiff(sit_com, com_com) %>% paste0(collapse = "|")

tec_sit_cae <- tibble::tibble(
    community = tec_sit$COM_ID,
    code=tec_sit$S_ID,
    name=tec_sit$S_ID,
    description = glue::glue("Site ID {tec_sit$S_ID} inside TEC Boundary {tec_sit$BDY_ID}"),
    area_type = 12, # 10 bound, 11 buffer, 12 site point
    source = 15, # TEC sit
    source_id = tec_sit$S_ID,
    encountered_on=default_date,
    encountered_by=1,
    point = tec_sit$geometry %>% geojsonsf::sfc_geojson(.) 
    ) #%>%  # (123560) dplyr::filter(!grepl(site_name_typos, community)) # 105606
```

TODO: exclude / QA community names
```{r tec_sit_tsc, eval=F}
# 7008 sites
tec_sit_cae %>% 
    dplyr::filter(!grepl(fat_filter, community)) %>%
    dplyr::filter(!grepl("Nullarbor microbial mantles", community)) %>% 
    chunk_post(serializer = "occ-community-points", api_url = dev, verbose = T)

# 10 sites, site shapefile mismatches com code
tec_sit_cae %>% 
  dplyr::filter(community == "Nullarbor microbial mantles") %>% 
  dplyr::mutate(community = "Microbial mantles") %>% 
  chunk_post(serializer = "occ-community-points", api_url = dev, verbose = T)

# 119 sites
tec_sit_cae %>% 
  dplyr::filter(community == "Limestone ridges (SCP 26a)") %>% 
  chunk_post(serializer = "occ-community-points", api_url = dev, verbose = T)

# 17k sites
# TODO update TSC community name of "Banksia Woodlands of the Swan Coastal Plain" (is None)
tec_sit_cae %>% 
  dplyr::filter(community == "Banksia Woodlands of the Swan Coastal Plain") %>% 
  chunk_post(serializer = "occ-community-points", api_url = dev, verbose = T)
```


### Long running data import
This will run for about a day.
```{r tec_import_long, eval=F}
tec_sit_cae %>% 
    dplyr::filter(grepl(fat_filter, community)) %>%
    chunk_post(serializer = "occ-community-points", api_url = prod, verbose = T)

tec_bdy_cae %>% 
    filter(grepl(fat_filter, community)) %>%
    chunk_post(serializer = "occ-community-areas", api_url = prod, verbose = T)

tec_buf_cae %>% 
    dplyr::filter(grepl(fat_filter, community)) %>%
    chunk_post(serializer = "occ-community-areas", api_url = prod, verbose = T)
```


### Occurrences
TEC refers to a single AOO as "occurrence". Occurrences may be geo-referenced as boundaries and/or
buffered boundaries (buffers) and/or through point sites within.

Occurrences refer to `BDI_ID` (boundary_id), which is a PK to a bunch of GIS tables.

```{r tec_occ}
# what's that?
classifications <- tec$CLASSIFICATIONS %>%
  dplyr::transmute(
    classification_code = CLASS.CODE %>% as.character(),
    classification = CLASS.DESC %>% as.character()
  )

# Occurrence confidentiality
# confidentiality <- tec$OCC_CONFIDENTIAL_CODES %>%
#   dplyr::transmute(
#     confidential_id = OCC.CONFIDENTIAL %>% as.integer(),
#     is_confidential = OCC.CONFID.DESC %>% as.character()
#   )

reliability <- tec$RELIABILITY %>% dplyr::transmute(
  reliability_id = BR.CODE %>% as.character(),
  reliability = BR.DESC %>% as.character()
)

tec_occ <- tec$OCCURRENCES %>%
  tibble::as.tibble() %>%
  dplyr::transmute(
    # IDs
    occ_id = OCC.UNIQUE.ID %>% as.integer(),
    com_id = COM.NO %>% as.integer(),
    occ_no = OCC.NO %>% as.integer(),
    source_code = OCC.SOURCE.CODE %>% as.character(),

    # spatial references
    boundary_id = BDY.ID %>% as.integer(),
    boundary_desc = OCC.BOUNDARY.DESC %>% as.character(),
    # boundary_no = OCC.NO.BOUNDARY %>% as.integer, # empty
    buffer_radius = OCC.BUFFER.RADIUS %>% as.integer(),
    original_area = OCC.ORIGINAL.AREA %>% as.double(),
    area_accuracy = OCC.AREA.ACCURACY %>% as.double(),

    # occ details
    species_desc = OCC.SPECIES.DESC %>% as.character(),
    commonwealth_desc = OCC.COMMONWEALTH.DESC %>% as.character(),
    confidential_id = OCC.CONFIDENTIAL %>% as.integer(),
    reliability_id = OCC.BR.CODE %>% as.character(),
    desc = OCC.DESC %>% as.character(),
    status_code = OCC.STATUS.CODE %>% as.logical(),
    ctrc_sys = OCC.CTRC.SYS %>% as.integer(), # FK CTRC_SYSTEMS
    ctrc_recommendation = OCC.CTRC.RECOMMENDATION %>% as.character(),
    cwealth_listing = OCC.CWEALTH.LISTING %>% as.character(),
    data = OCC.DATA %>% as.character(),
    other = OCC.OTHER %>% as.character(),
    soil = OCC.SOIL %>% as.character(),
    surf_geology = OCC.SURF.GEOLOGY %>% as.character(),
    land_element = OCC.LAND.ELEMENT %>% as.character(),
    water = OCC.WATER %>% as.character(),
    drainage = OCC.DRAINAGE %>% as.character(),
    com_structure = OCC.COM.STRUCTURE %>% as.character(),
    classification = OCC.CLASSIFICATION %>% as.character(),
    other_attr = OCC.OTHER.ATTR %>% as.character(),

    # spatial lookups
    beard_dist_code = OCC.BEARD.DIST.CODE %>% as.character(),
    beard_map_code = OCC.BEARD.MAP.CODE %>% as.character(),
    beard_desc = OCC.BEARD.DESC %>% as.character(),
    ibra_reg_code = OCC.IBRA.REG.CODE %>% as.character(),
    dola_ref = OCC.DOLA.REF %>% as.character(),
    bush_forever_site_no = OCC.BUSH.FOREVER.SITE.NO %>% as.integer(),
    zone_code = OCC.ZONE.CODE %>% as.character(),

    # audit
    created_by = USERNAME %>% as.character(),
    created_on = OCC.DATE.ENTERED %>%
      lubridate::parse_date_time(., orders = orders, tz = tz),
    changed_on = OCC.DATE.EDITED %>%
      as.character() %>%
      lubridate::parse_date_time(., orders = orders, tz = tz)
  ) %>%
  dplyr::left_join(com_short, by = "com_id") %>%
  # dplyr::left_join(confidentiality, by = "confidential_id") %>%
  dplyr::left_join(reliability, by = "reliability_id") %>%
  dplyr::select(-confidential_id, reliability_id)
tec_occ %>% skimr::skim()
tec_occ %>% head(100) %>% DT::datatable(.)

# unique(occurrences$ctrc_sys) %>% sort()

occ_short <- tec_occ %>%
  dplyr::select(occ_id, boundary_id, boundary_desc, com_id, com_label, com_name)

# tec_occ %>% dplyr::filter(com_label == "SCP20a") %>% View

```

## Occurrence Actions

```{r occ_actions}
occ_actions <- tec$OCCURRENCE_ACTIONS %>%
  dplyr::transmute(
    occ_id = OCC.UNIQUE.ID %>% as.integer(),
    occ_action_id = OACT.NO %>% as.integer(),
    occ_action_date = OACT.DATE %>% parse_date_time(., orders = orders, tz = tz),
    occ_action_description = OACT.DESC %>% as.character(),
    username = USERNAME %>% as.character(),
    action_id = OACT.ACTION.CODE %>% as.character()
  ) %>%
  dplyr::left_join(occ_short, by = "occ_id") %>%
  dplyr::left_join(actions, by = "action_id")

occ_actions %>% DT::datatable(.)
```

## Occcurrence Recommendations
```{r occ_recommendations}
occ_recommendations <- tec$OCCURRENCE_RECOMMENDATIONS %>%
  dplyr::transmute(
    occ_id = OCC.UNIQUE.ID %>% as.integer(),
    rank = OREC.RANK %>% as.character(),
    recommended_on = OREC.DATE %>% as.character(),
    recommendation_code = OREC.REC.CODE %>% as.character(),
    recommendation = OREC.DESC %>% as.character(),
    username = USERNAME %>% as.character()
  ) %>%
  dplyr::left_join(occ_short, by = "occ_id") %>%
  dplyr::left_join(recommendations, by = "recommendation_code")

occ_recommendations %>% DT::datatable(.)
```

## Occurrence Extent

```{r occ_extent}
sources <- tec$SOURCES %>%
  dplyr::transmute(
    source_code = SOURCE.CODE %>% as.character(),
    source = SOURCE.DESC %>% as.character()
  )

occ_extent <- tec$EXTENTS %>%
  dplyr::transmute(
    occ_id = OCC.UNIQUE.ID %>% as.integer(),
    EXT.NO = EXT.NO %>% as.character(),
    EXT.DATE = EXT.DATE %>% as.character(),
    source_code = EXT.SOURCE.CODE %>% as.character(),
    EXT.AREA = EXT.AREA %>% as.character(),
    EXT.AREA.ACC = EXT.AREA.ACC %>% as.character(),
    area_decline_id = EXT.AD.CODE %>% as.integer(),
    EXT.COMMENT = EXT.COMMENT %>% as.character(),
    USERNAME = USERNAME %>% as.character()
  ) %>%
  dplyr::left_join(occ_short, by = "occ_id") %>%
  dplyr::left_join(sources, by = "source_code") %>%
  dplyr::left_join(area_decline, by = "area_decline_id")

occ_extent %>% DT::datatable(.)
```

## Additional Data

```{r occ_additional_data}
items <- tec$ITEMS %>%
  dplyr::transmute(
    item_code = ITEM.CODE %>% as.character(),
    item = ITEM.DESC %>% as.character()
  )

occ_additional_data <- tec$ADDITIONAL_DATA %>%
  dplyr::transmute(
    occ_id = OCC.UNIQUE.ID %>% as.integer(),
    additional_data_id = ADD.NO %>% as.integer(),
    item_code = ADD.ITEM.CODE %>% as.character(),
    description = ADD.DESC %>% as.character(),
    username = USERNAME %>% as.character()
  ) %>%
  dplyr::left_join(occ_short, by = "occ_id") %>%
  dplyr::left_join(items, by = "item_code")

occ_additional_data %>% DT::datatable(.)
```

## Occcurrence Maps
Names of physical maps describing one or multiple occurrences.

```{r occ_maps}
maps <- tec$MAPS
occ_maps <- tec$OCCURRENCE_MAPS
```

## Biological Processes
```{r occ_biol_process}
occ_biol_process <- tec$BIOL_PROCESSES %>%
  dplyr::transmute(
    occ_id = OCC.UNIQUE.ID %>% as.integer(),
    biol_process_id = BIOL.NO %>% as.integer(),
    biol_process = BIOL.PROCESS %>% as.character(),
    username = USERNAME %>% as.character()
  ) %>%
  dplyr::left_join(occ_short, by = "occ_id")

occ_biol_process %>% DT::datatable(.)
```

## Non-biological Processes
```{r occ_nonbiol_process}
occ_nonbiol_process <- tec$NON_BIOL_PROCESSES %>%
  dplyr::transmute(
    occ_id = OCC.UNIQUE.ID %>% as.integer(),
    nonbiol_process_id = NBIO.NO %>% as.integer(),
    nonbiol_process = NBIO.PROCESS %>% as.character(),
    username = USERNAME %>% as.character()
  ) %>%
  dplyr::left_join(occ_short, by = "occ_id")

occ_nonbiol_process %>% DT::datatable(.)
```

## Fire History

```{r occ_fire_history}
fire_history <- tec$FIRE_HISTORY %>%
  dplyr::transmute(
    occ_id = OCC.UNIQUE.ID %>% as.integer(),
    fire_history_id = FIRE.NO %>% as.integer(),
    fire_history_date = FIRE.DATE %>% as.character(),
    fire_history_comment = FIRE.COMMENT %>% as.character(),
    username = USERNAME %>% as.character()
  ) %>%
  dplyr::left_join(occ_short, by = "occ_id")

fire_history %>% DT::datatable(.)
```

## Fauna Occurrences
There is another separate group of occurrences:

* Occurrence Species
* Occurrence Species INI (contains an external db config)
* Occurrence Fauna
* Species Roles
* Fauna Taxa
* Fauna Groups
* Fauna Conservation Codes

What is the overlap to the Threatened Fauna DB in terms of meaning and data flow?

```{r}
# lookups
fauna_conservation_codes <- tec$FAUNA_CONSERVATION_CODES %>%
  dplyr::transmute(
    conservation_code = CONSV.CODE %>% as.character(),
    conservation_code_description = CONSV.DESC %>% as.character()
  )

fauna_groups <- tec$FAUNA_GROUPS # paraphyletic groups

species_roles <- tec$SPECIES_ROLES %>%
  dplyr::transmute(
    species_role_code = SP.ROLE.CODE %>% as.character(),
    species_role = SP.ROLE.DESC %>% as.character()
  )

fauna_taxa <- tec$FAUNA_TAXA # taxon_id

# primary
occ_fauna <- tec$OCCURRENCE_FAUNA %>%
  dplyr::transmute(
    occ_id = OCC.UNIQUE.ID %>% as.integer(),
    taxon_id = OF.TAXON.ID %>% as.character(),
    species_code = OF.SPCODE %>% as.character(),
    species_role_code = OF.SP.ROLE.CODE %>% as.character(),
    comments = OF.VOUCHER.NO %>% as.character(),
    username = USERNAME %>% as.character(),
  ) %>%
  dplyr::left_join(occ_short, by = "occ_id") %>%
  dplyr::left_join(species_roles, by = "species_role_code")

occ_fauna %>% DT::datatable(.)

occ_species <- tec$OCCURRENCE_SPECIES %>%
  dplyr::transmute(
    occ_id = OCC.UNIQUE.ID %>% as.integer(),
    taxon_id = SPEC.TAXON.ID %>% as.character(),
    species_code = SPEC.SPCODE %>% as.character(),
    species_role_code = SPEC.SP.ROLE.CODE %>% as.character(),
    comments = SPEC.VOUCHER.NO %>% as.character(),
    username = USERNAME %>% as.character(),
  ) %>%
  dplyr::left_join(occ_short, by = "occ_id") %>%
  dplyr::left_join(species_roles, by = "species_role_code")

# occ_species %>% DT::datatable(.)
```

# Surveys
`Surveys` are full site visits with a complete set of measurements.
`Site visits` are quick site visits with a partial set of measurements.
Both relate to Occurrences.

A separate group related to surveys has the tables:

* Threat Impacts
* Threats
* Surveys
* Conditions
* Survey Conditions
* Survey Threats

```{r surveys}
# lookups
impacts <- tec$THREAT_IMPACTS %>%
  dplyr::transmute(
    impact_code = IMP.CODE %>% as.character(),
    impact = IMP.DESC %>% as.character()
  )

threats <- tec$THREATS %>%
  dplyr::transmute(
    threat_code = THREAT.CODE %>% as.character(),
    threat = THREAT.DESC %>% as.character()
  )

conditions <- tec$CONDITIONS %>%
  dplyr::transmute(
    condition_code = COND.CODE %>% as.character(),
    condition = COND.DESC %>% as.character()
  )

# primary
com_threats <- tec$COMMUNITY_THREATS %>%
  dplyr::transmute(
    com_id = COM.NO %>% as.integer(),
    id = CTHR.NO %>% as.character(),
    date = CTHR.DATE %>% as.character(),
    threat_code = CTHR.THREAT.CODE %>% as.character(),
    description = CTHR.DESC %>% as.character()
  ) %>%
  dplyr::left_join(threats, by = "threat_code")

surveys <- tec$SURVEYS %>%
  dplyr::transmute(
    occ_id = OCC.UNIQUE.ID %>% as.integer(),
    survey_id = SUR.NO %>% as.integer(),
    surveyed_on = SUR.DATE %>% as.character(),
    surveyed_by = SUR.SURVEYOR %>% as.character(),
    survey_comments = SUR.COMMENTS %>% as.character(),
    username = USERNAME %>% as.character()
  ) %>%
  dplyr::left_join(occ_short, by = "occ_id")

survey_threats <- tec$SURVEY_THREATS %>%
  dplyr::transmute(
    occ_id = OCC.UNIQUE.ID %>% as.integer(),
    survey_id = SUR.NO %>% as.integer(),
    threat_rank = THR.RANK %>% as.integer(),
    threat_code = THR.THREAT.CODE %>% as.character(),
    threat_percent = THR.PERCENT %>% as.double(),
    historic_impact_code = THR.HIST.IMP.CODE %>% as.character(),
    potential_impact_code = THR.POT.IMP.CODE %>% as.character(),
    modification = THR.MODIFICATION %>% as.character(),
    comments = THR.COMMENTS %>% as.character(),
    username = USERNAME %>% as.character(),
  ) %>%
  dplyr::left_join(occ_short, by = "occ_id") %>%
  dplyr::left_join(threats, by = "threat_code") %>%
  dplyr::left_join(impacts, by = c("historic_impact_code" = "impact_code")) %>%
  dplyr::rename(historic_impact = impact) %>%
  dplyr::left_join(impacts, by = c("potential_impact_code" = "impact_code")) %>%
  dplyr::rename(potential_impact = impact)

survey_conditions <- tec$SURVEY_CONDITIONS %>%
  dplyr::transmute(
    occ_id = OCC.UNIQUE.ID %>% as.integer(),
    survey_id = SUR.NO %>% as.integer(),
    condition_code = SCON.COND.CODE %>% as.character(),
    condition_percent = SCON.PERCENT %>% as.double(),
    comments = SCON.COMMENTS %>% as.character(),
    username = USERNAME %>% as.character()
  ) %>%
  dplyr::left_join(occ_short, by = "occ_id") %>%
  dplyr::left_join(conditions, by = "condition_code")
```

Out of `r nrow(surveys)` surveys, there are only `r length(unique(surveys$survey_id))` distinct `survey_id`s.
With `length(unique(surveys$surveyed_on))`, this means that a `survey_id` does not denote an individual
visit of a TEC boundary.

However, there are exactly as many unique combinations of `survey_id` and `occ_id` 
(`r surveys %>% group_by(occ_id, survey_id) %>% tally() %>% nrow`) as records in the table `surveys`.

How do `r nrow(survey_conditions)` `survey_conditions` reference `r nrow(surveys)` `surveys`? Likely by `occ_id` and `survey_id`.

# Site Visits
* Sites: Point-referenced sites, mostly 10x10m or 6x6m quadrats, within boundaries.
* Datum: coordinate datum
* Site Visits: CommunityAreaEncounters of Sites (lat/lon point) at date (site_visit$site_visit_date).
* Site Species: Species encountered during a site_visit.

```{r sites}
datum <- tec$DATUM %>%
  dplyr::transmute(
    datum_code = DATUM.CODE %>% as.character(),
    datum = DATUM.DESC %>% as.character(),
    semi_major_axis = SEMI.MAJOR.AXIS %>% as.double(),
    inverse_flattening = INVERSE.FLATTENING %>% as.double()
  )

sites <- tec$SITES %>%
  dplyr::transmute(
    occ_id = OCC.UNIQUE.ID %>% as.integer(),
    site_code = S.ID %>% as.character(),
    datum_code = S.DATUM.CODE %>% as.character(),
    latitude = S.LATITUDE %>% as.double(),
    longitude = S.LONGITUDE %>% as.double(),
    comments = S.COMMENTS %>% as.character(),
    # S.LATITUDE.PREF = S.LATITUDE.PREF  %>% as.character, # wtf?
    # S.LONGITUDE.PREF = S.LONGITUDE.PREF  %>% as.character, # wut?
    created_by = USERNAME %>% as.character(),
    changed_on = S.DATE.EDITED %>% parse_date_time(., orders = orders, tz = tz),
    permanently_marked = S.PERMANENTLY.MARKED %>% as.character(), # "", "Y"
    location = S.LOCATION %>% as.character(),
    habitat = S.HABITAT %>% as.character(),
    soil = S.SOIL %>% as.character(),
    # rock_type = S.ROCK.TYPE %>% as.character(), # empty
    # parent_material = S.PARENT.MATERIAL %>% as.character(), # empty
    vegetation = S.VEGETATION %>% as.character(),
    # rock_pile_notes = S.ROCK.PILE.NOTES %>% as.character(), # empty
    site_notes = S.SITE.NOTES %>% as.character()
  ) %>%
  dplyr::left_join(occ_short, by = "occ_id") %>%
  dplyr::left_join(datum, by = "datum_code")

# sites[15000:15250,] %>% skim # CKAN datapusher chokes on this chunk

site_visit <- tec$SITE_VISITS %>%
  dplyr::transmute(
    site_visit_id = SITE.VISIT.ID %>% as.integer(),
    site_code = S.ID %>% as.character(),
    site_visit_date = SV.VISIT.DATE %>% as.character(),
    project_code = SV.PROJ.CODE %>% as.character(),
    site_visit_described_by = SV.DESCRIBED.BY %>% as.character(),
    site_visit_described_by_notes = SV.DESCRIBED.BY.NOTES %>% as.character(),
    seasonal_conditions = SV.SEASONAL.CONDITIONS %>% as.character(),
    observation_type = SV.OBSERVATION.TYPE %>% as.character(),
    observation_quality = SV.OBSERVATION.QUALITY %>% as.character(),
    plant_id_quality = SV.PLANT.ID.QUALITY %>% as.character(),
    site_dimensions = SV.SITE.DIMENSIONS %>% as.character(),
    site_uniformity = SV.SITE.UNIFORMITY %>% as.character(),
    site_status = SV.SITE.STATUS %>% as.character(),
    photo_numbers = SV.PHOTO %>% as.character(),
    vegetation_condition = SV.VEGETATION.CONDITION %>% as.character(),
    fire_age = SV.FIRE.AGE %>% as.character(),
    fire_notes = SV.FIRE.NOTES %>% as.character(),
    observation_notes = SV.OBSERVATION.NOTES %>% as.character()
  ) %>% 
  dplyr::left_join(sites, by = "site_code")

# coll code and number indicates voucher specimens
voucher_specimens <- tec$SITE_SPECIES %>%
  dplyr::transmute(
    site_species_id = SITE.SPECIES.ID %>% as.character(),
    site_visit_id = SITE.VISIT.ID %>% as.integer(),
    name_id = SSP.NAME.ID %>% as.character(),
    species_code = SSP.SPCODE %>% as.character(),
    height = SSP.HEIGHT %>% as.character(),
    collector_code = SSP.COLLECTOR.CODE %>% as.character(),
    collection_number = SSP.COLLECTION.NUMBER %>% as.character(),
    notes = SSP.NOTES %>% as.character()
  ) %>%
  dplyr::left_join(site_visit, by = "site_visit_id")
```

There are `r nrow(sites)` point locations known as TEC sites.

```{r sites_map, fig.height=5, fig.width=7}
leaflet(width = 800, height = 600) %>%
  addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
  addProviderTiles("OpenStreetMap.Mapnik", group = "Place names") %>%
  clearBounds() %>%
  addAwesomeMarkers(
    data = sites,
    lng = ~ longitude, lat = ~ latitude,
    icon = leaflet::makeAwesomeIcon(icon = "leaf"),
    label = ~ paste(site_code),
    popup = ~ paste(
      "<h3>Site", site_code, "</h3>",
      "<strong>Community</strong>", com_label, "<br/>",
      "<strong>Created by</strong>", created_by, "on", changed_on, "<br/>",
      "<strong>Comments</strong>", comments, "<br/>",
      "<strong>Permanent</strong>", permanently_marked, "<br/>",
      "<strong>Location</strong>", location, "<br/>",
      "<strong>Habitat</strong>", habitat, "<br/>",
      "<strong>Soil</strong>", soil, "<br/>",
      "<strong>Vegetation</strong>", vegetation, "<br/>",
      "<strong>Site notes</strong>", site_notes, "<br/>"
    ),
    group = "Site points",
    clusterOptions = markerClusterOptions()
  ) %>%
  addLayersControl(
    baseGroups = c("Aerial", "Place names"),
    overlayGroups = c("Site points"),
    options = layersControlOptions(collapsed = FALSE)
  )
```


# Gazettal
The data model of gazettals requires clarification.

```{r gazettal}
# lookups
library(magrittr)
tec_cons_categories <- tec$CATEGORY_TYPES %>%
  dplyr::transmute(
    category_code = CT.TYPE %>% as.character(),
    category_label = CT.DESC %>% as.character(),
    authority = CT.AUTHORITY.CODE %>% as.character()
  )

tec_cons_criteria <- tec$CRITERIA_CODES %>%
  dplyr::transmute(
    criterion_code = CRIT.CODE %>% as.character(),
    criterion_description = CRIT.DESC %>% as.character()
  )

# Using a built-in function to turn uppercase into lowercase? Nah. Hold my beer.
authorities <- tec$AUTHORITIES # A lookup for upper cased authorities vs lower cased authorities.
endorsement <- tec$ENDORSED_BY # another lookup for lower case.

# primary
tec_cons_listing_categories <- tec$CATEGORIES %>%
  dplyr::transmute(
    com_id = COM.NO %>% as.integer(),
    effective_from = CAT.EFFECT.DATE %>% as.character(),
    category_code = CAT.CT.TYPE %>% as.character(),
    cat_comment = CAT.COMMENT %>% as.character(),
    cat_review_date = CAT.REVIEW.DATE %>% as.character(),
    cat_endorsed_code = CAT.ENDORSED.CODE %>% as.character(),
    cat_endorsed_date = CAT.ENDORSED.DATE %>% as.character(),
    cat_endorsed_by_minister = CAT.ENDORSED.BY.MINISTER %>% as.character()
  ) %>%
  dplyr::left_join(com_short, by = "com_id") %>%
  dplyr::left_join(tec_cons_categories, by = "category_code")

tec_cons_listing_criteria <- tec$CATEGORY_CRITERIA %>%
  dplyr::transmute(
    com_id = COM.NO %>% as.integer(),
    effective_from = CAT.EFFECT.DATE %>% as.character(),
    criterion_code = CC.CRIT.CODE %>% as.character()
  ) %>%
  dplyr::left_join(com_short, by = "com_id") %>%
  dplyr::left_join(tec_cons_criteria, by = "criterion_code")
# join com_short, criteria codes
```

# Recovery Plans

```{r recovery_plans}
recovery_plans <- tec$INTERIM_RECOVERY_PLANS %>%
  dplyr::transmute(
    recovery_plan_id = IRP.NO %>% as.character(),
    com_id = COM.NO %>% as.integer(),
    endorsed_on = IRP.DATE.ENDORSED %>% parse_date_time(., orders = orders, tz = tz),
    ends_on = IRP.END.DATE %>% parse_date_time(., orders = orders, tz = tz),
    author = IRP.AUTHOR %>% as.character(),
    reviewed_on = IRP.DATE.REVIEWED %>% parse_date_time(., orders = orders, tz = tz),
    comments = IRP.COMMENTS %>% as.character()
  ) %>%
  dplyr::left_join(com_short, by = "com_id")
# make_coldef(recovery_plans)
recovery_plans %>% skim()
```

# Data upload
Create resource once-off to get a resource ID.
Also upload the compiled workbook.
Note, this process lags one compilation of this workbook, as the HTML output
is generated after the R code is run - knit a workbook twice to update the data catalogue.

```{r tec_upload}
d <- ckanr::package_show("threatened-ecological-communities-database")
# ckanr::resource_update("80009375-19e2-45d3-a1ae-068143fa03a5", "data_etl_tec.html") # HTML, not CSV
upload_to_ckan(communities, "Threatened Ecological Communities (TEC)", 
               d$id, resource_id = "9782cc52-7be8-494a-a7da-51845c119a21")
upload_to_ckan(tec_occ, "TEC Occurrences", 
               d$id, resource_id = "65f43392-a8c5-41e2-a31e-cc4135e9274e")
upload_to_ckan(sites, "TEC Site Points", 
               d$id, resource_id = "bf4c2973-e549-408b-a583-013052072d6c")
upload_to_ckan(com_recommendations, "TEC Recommendations", 
               d$id, resource_id = "fd96a053-523c-401d-8755-9116c55f5744")
upload_to_ckan(com_publications, "TEC Publications", 
               d$id, resource_id = "fa51cd96-864a-4c43-a0e1-fc24621bb628")
upload_to_ckan(com_reviews, "TEC Reviews", 
               d$id, resource_id = "7c39b584-bc4b-4234-aff7-ddf4808692ed")
upload_to_ckan(com_actions, "TEC Actions", 
               d$id, resource_id = "17d3c4e2-df98-4ff6-9d4d-99ba54808bb4")
upload_to_ckan(com_history, "TEC History", 
               d$id, resource_id = "479d7358-d42d-4c4b-93ee-1bac39d21024")
upload_to_ckan(occ_actions, "TEC Occurrence Actions", 
               d$id, resource_id = "25894a68-11de-4a58-b521-6a9abd06ee40")
upload_to_ckan(occ_recommendations, "TEC Occurrence Recommendations", 
               d$id, resource_id = "747464e4-2434-459a-b654-52861b37095c")
upload_to_ckan(occ_extent, "TEC occurrence extent", 
               d$id, resource_id = "502c74d7-32be-453f-aff6-c50aedd3deed")
upload_to_ckan(occ_additional_data, "TEC occurrence additional data", 
               d$id, resource_id = "8dd3454c-1bf0-4575-adcb-154b700c4283")
upload_to_ckan(occ_biol_process, "TEC biological process", 
               d$id, resource_id = "5805a671-576a-4077-b595-35ca96d7ca96")
upload_to_ckan(occ_nonbiol_process, "TEC nonbiological process", 
               d$id, resource_id = "a5341117-2735-4694-8cd4-e806f4c06668")
upload_to_ckan(fire_history, "TEC fire history", 
               d$id, resource_id = "e062f57e-f6aa-4403-a682-875dab2f77f6")
upload_to_ckan(occ_fauna, "TEC occurence fauna", 
               d$id, resource_id = "a0eddc66-5fef-4252-b49f-691422a88f28")
upload_to_ckan(occ_species, "TEC occurence species", 
               d$id, resource_id = "e94f1e35-30da-4982-b851-ceb888eb41b3")
upload_to_ckan(com_threats, "TEC community threats", 
               d$id, resource_id = "5426c45b-1dee-4547-a6e3-bec81585e52e")
upload_to_ckan(survey_threats, "TEC survey threats", 
               d$id, resource_id = "5b795656-a240-402e-92b8-edbe54b3fee8")
upload_to_ckan(site_visit, "TEC site visit", 
               d$id, resource_id = "85cef94b-d12e-4aff-bd24-446d1c0757db")
upload_to_ckan(voucher_specimens, "TEC voucher specimens", 
               d$id, resource_id = "6b8fe493-8183-4fe8-a91a-67a569bf9c46")
upload_to_ckan(tec_cons_listing_categories, "TEC Conservation Listing Categories", 
               d$id, resource_id = "f0085651-8c38-496e-a6bd-1adc7ae2f06a")
upload_to_ckan(tec_cons_listing_criteria, "TEC Conservation Listing Criteria", 
               d$id, resource_id = "9483f333-97fd-4975-a799-d44bf8bd252a")
```


