---
title: 'Data ETL: Flora'
author: "Florian Mayer and Milly Piggott, DBCA"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE)
source("helpers.R")
```

# Context
This workbook summarises the legacy data migration and QA for the Threatened
and Priority Flora database (TFL).


TODO diagram data flow

By default, all code blocks are collapsed to provide better readability to the
non-technical audience. Feel free to expand the code blocks to view the process.

The code for all SCB workbooks is under version control at 
[github](https://github.com/dbca-wa/scarab-scripts).


# Extract and transform legacy Data from TFL
Data from the Threatened Flora Database (TFL) is downloaded from the 
[TFL dataset](https://data.dpaw.wa.gov.au/dataset/threatened-and-priority-flora-database) 
on the DBCA data catalogue, then extracted and transformed.

```{r load_tfl, message=FALSE}
tfl_data <- dl_mdbzip("77062fa0-f409-4fe7-ae28-de8c933a59bd")

# Look up tables
populations <- tfl_data$DRF_POPULATION %>%
  dplyr::transmute(
    pop_id = POP.ID %>% as.character(),
    pop_number = POP.NUMBER %>% as.character(),
    subpop_code = SUBPOP.CODE %>% as.character(),
    taxonid = TAXONID %>% as.character(),
    spname = SPNAME %>% as.character(),
    parent_pop_id = PARENT.POP.ID %>% as.character(),
    pop_rec_type_ind = POP.REC.TYPE.IND %>% as.character(),
    creation_sheetno = CREATION.SHEETNO %>% as.character(),
    first_observed_date = FIRST.OBSERVED.DATE %>% as.character(),
    confirmed = CONFIRMED %>% as.character(),
    status = STATUS %>% as.character(),
    pop_comments = POP.COMMENTS %>% as.character(),
    assoc_sp_accum = ASSOC.SP.ACCUM %>% as.character(),
    location = LOCATION %>% as.character(),
    district = DISTRICT %>% as.character(),
    landdistrict = LANDDISTRICT %>% as.character(),
    lga_code = LGA.CODE %>% as.character(),
    nearplace = NEARPLACE %>% as.character(),
    distance = DISTANCE %>% as.character(),
    direction = DIRECTION %>% as.character(),
    purpose1 = PURPOSE1 %>% as.character(),
    purpose2 = PURPOSE2 %>% as.character(),
    vesting = VESTING %>% as.character(),
    land_loc_no = LAND.LOC.NO %>% as.character(),
    reserve = RESERVE %>% as.character(),
    class = CLASS %>% as.character(),
    utility_ref = UTILITY.REF %>% as.character(),
    land_manager = LAND.MANAGER %>% as.character(),
    land_mgr_address = LAND.MGR.ADDRESS %>% as.character(),
    land_mgr_phone = LAND.MGR.PHONE %>% as.character(),
    co_ord_type_code = CO.ORD.TYPE.CODE %>% as.character(),
    co_ord_source_code = CO.ORD.SOURCE.CODE %>% as.character(),
    gda94lat = GDA94LAT %>% as.character(),
    gda94long = GDA94LONG %>% as.character(),
    loc_sect_date = LOC.SECT.DATE %>% as.character(),
    datum = DATUM %>% as.character(),
    latdeg = LATDEG %>% as.character(),
    latmin = LATMIN %>% as.character(),
    latsec = LATSEC %>% as.character(),
    longdeg = LONGDEG %>% as.character(),
    longmin = LONGMIN %>% as.character(),
    longsec = LONGSEC %>% as.character(),
    gps_zone = GPS.ZONE %>% as.character(),
    northing = NORTHING %>% as.character(),
    easting = EASTING %>% as.character(),
    no_of_satellites = NO.OF.SATELLITES %>% as.character(),
    resolution = RESOLUTION %>% as.character(),
    map_used = MAP.USED %>% as.character(),
    map_scale = MAP.SCALE %>% as.character(),
    dec_latitude = DEC.LATITUDE %>% as.character(),
    dec_longitude = DEC.LONGITUDE %>% as.character(),
    count_sect_date = COUNT.SECT.DATE %>% as.character(),
    svy_extent = SVY.EXTENT %>% as.character(),
    svy_effort_area = SVY.EFFORT.AREA %>% as.character(),
    svy_effort_time = SVY.EFFORT.TIME %>% as.character(),
    svy_effort = SVY.EFFORT %>% as.character(),
    quality_ind = QUALITY.IND %>% as.character(),
    count_mthd_code = COUNT.MTHD.CODE %>% as.character(),
    mature_plants = MATURE.PLANTS %>% as.character(),
    mature_dead = MATURE.DEAD %>% as.character(),
    juvenile_plants = JUVENILE.PLANTS %>% as.character(),
    juvenile_dead = JUVENILE.DEAD %>% as.character(),
    seedling_plants = SEEDLING.PLANTS %>% as.character(),
    seedling_dead = SEEDLING.DEAD %>% as.character(),
    cnt_plant_type_code = CNT.PLANT.TYPE.CODE %>% as.character(),
    simple_live_tot = SIMPLE.LIVE.TOT %>% as.character(),
    area_occupied = AREA.OCCUPIED %>% as.character(),
    area_occupied_method = AREA.OCCUPIED.METHOD %>% as.character(),
    quad_num = QUAD.NUM %>% as.character(),
    quad_size = QUAD.SIZE %>% as.character(),
    quad_tot_sq_m = QUAD.TOT.SQ.M %>% as.character(),
    quad_num_mature = QUAD.NUM.MATURE %>% as.character(),
    quad_num_juvenile = QUAD.NUM.JUVENILE %>% as.character(),
    quad_num_seedlings = QUAD.NUM.SEEDLINGS %>% as.character(),
    quad_num_total = QUAD.NUM.TOTAL %>% as.character(),
    clonal = CLONAL %>% as.character(),
    vegetative = VEGETATIVE %>% as.character(),
    in_buds = IN.BUDS %>% as.character(),
    in_flower = IN.FLOWER %>% as.character(),
    immature_fruit = IMMATURE.FRUIT %>% as.character(),
    fruit = FRUIT %>% as.character(),
    dehisced_fruit = DEHISCED.FRUIT %>% as.character(),
    flower_percentage = FLOWER.PERCENTAGE %>% as.character(),
    hab_sect_date = HAB.SECT.DATE %>% as.character(),
    population_condition = POPULATION.CONDITION %>% as.character(),
    population_notes = POPULATION.NOTES %>% as.character(),
    habitat_condition = HABITAT.CONDITION %>% as.character(),
    habitat_notes = HABITAT.NOTES %>% as.character(),
    soil_condition = SOIL.CONDITION %>% as.character(),
    landform = LANDFORM %>% as.character(),
    rock_type = ROCK.TYPE %>% as.character(),
    gravel = GRAVEL %>% as.character(),
    soil_type = SOIL.TYPE %>% as.character(),
    soil_color = SOIL.COLOR %>% as.character(),
    drainage = DRAINAGE %>% as.character(),
    aspect = ASPECT %>% as.character(),
    associated_species = ASSOCIATED.SPECIES %>% as.character(),
    fire_intensity = FIRE.INTENSITY %>% as.character(),
    fire_year = FIRE.YEAR %>% as.character(),
    fire_season = FIRE.SEASON %>% as.character(),
    fencing_status = FENCING.STATUS %>% as.character(),
    fencing_comments = FENCING.COMMENTS %>% as.character(),
    roadside_marker_status = ROADSIDE.MARKER.STATUS %>% as.character(),
    rdside_mkr_comments = RDSIDE.MKR.COMMENTS %>% as.character(),
    other_comments = OTHER.COMMENTS %>% as.character(),
    vchr_sect_date = VCHR.SECT.DATE %>% as.character(),
    voucher_location = VOUCHER.LOCATION %>% as.character(),
    dupvouch_location = DUPVOUCH.LOCATION %>% as.character(),
    collector_no = COLLECTOR.NO %>% as.character(),
    barcode = BARCODE %>% as.character(),
    vchr_status_code = VCHR.STATUS.CODE %>% as.character(),
    active_ind = ACTIVE.IND %>% as.character(),
    deactivated_date = DEACTIVATED.DATE %>% as.character(),
    reason_deactivated = REASON.DEACTIVATED %>% as.character(),
    created_by = CREATED.BY %>% as.character(),
    created_date = CREATED.DATE %>% as.character(),
    modified_by = MODIFIED.BY %>% as.character(),
    modified_date = MODIFIED.DATE %>% as.character(),
    simple_dead_tot = SIMPLE.DEAD.TOT %>% as.character(),
    observer_code = OBSERVER.CODE %>% as.character(),
    obs_name = OBS.NAME %>% as.character(),
    observer_phone = OBSERVER.PHONE %>% as.character(),
    obs_role_code = OBS.ROLE.CODE %>% as.character(),
    obs_org_code = OBS.ORG.CODE %>% as.character(),
    obs_org_name = OBS.ORG.NAME %>% as.character(),
    land_mgr_notes = LAND.MGR.NOTES %>% as.character()
  )

pop_forms <- tfl_data$DRF_POP_FORMS %>%
  dplyr::transmute(
    pop_forms_id = POP.FORMS.ID %>% as.character(),
    pop_id = POP.ID %>% as.character(),
    sheetno = SHEETNO %>% as.character(),
    prev_pop_id = PREV.POP.ID %>% as.character(),
    active_ind = ACTIVE.IND %>% as.character(),
    deactivated_date = DEACTIVATED.DATE %>% as.character(),
    reason_deact = REASON.DEACT %>% as.character(),
    created_by = CREATED.BY %>% as.character(),
    created_date = CREATED.DATE %>% as.character(),
    current_rec = CURRENT.REC %>% as.character(),
    chg_type_code = CHG.TYPE.CODE %>% as.character(),
    deact_by = DEACT.BY %>% as.character()
  )
pop_forms[1:100, ] %>% DT::datatable(.)

pop_threats <- tfl_data$DRF_POP_THREATS %>%
  dplyr::transmute(
    pop_threat_id = POP.THREAT.ID %>% as.character(),
    pop_id = POP.ID %>% as.character(),
    threat_code = THREAT.CODE %>% as.character(),
    cur_impact = CUR.IMPACT %>% as.character(),
    pot_impact = POT.IMPACT %>% as.character(),
    onset = ONSET %>% as.character(),
    agent_code = AGENT.CODE %>% as.character(),
    comments = COMMENTS %>% as.character(),
    status = STATUS %>% as.character(),
    active_ind = ACTIVE.IND %>% as.character(),
    deactivated_date = DEACTIVATED.DATE %>% as.character(),
    reason_deactivated = REASON.DEACTIVATED %>% as.character(),
    created_by = CREATED.BY %>% as.character(),
    created_date = CREATED.DATE %>% as.character(),
    modified_by = MODIFIED.BY %>% as.character(),
    modified_date = MODIFIED.DATE %>% as.character(),
    sheetno = SHEETNO %>% as.character(),
    observation_date = OBSERVATION.DATE %>% as.character()
  )
pop_threats[1:100, ] %>% DT::datatable(.)

pop_vegclass <- tfl_data$DRF_POP_VEG_CLASSES %>%
  dplyr::transmute(
    veg_class_id = VEG.CLASS.ID %>% as.character(),
    veg_class_code = VEG.CLASS.CODE %>% as.character(),
    pop_id = POP.ID %>% as.character(),
    veg_class_dom1 = VEG.CLASS.DOM1 %>% as.character(),
    veg_class_dom2 = VEG.CLASS.DOM2 %>% as.character()
  )
pop_vegclass[1:100, ] %>% DT::datatable(.)

tfl_cons_listing <- tfl_data$DRF_TAXON_CONSV_LISTINGS %>%
  dplyr::transmute(
    txn_list_id = TXN.LIST.ID %>% as.character(),
    name_id = TAXONID %>% as.integer(),
    name = NAME %>% as.character(),
    list_code = LIST.CODE %>% as.character(),
    category_code = STATUS.CODE %>% as.character(),
    list_code_wa = WA.IUCN.RANK.LIST.CODE %>% as.character(),
    category_code_wa = WA.IUCN.RANK %>% as.character(),
    criteria_code_wa = IUCN.CRITERIA %>% as.character(),
    date_listed = DATE.LISTED %>% 
      parse_date_time(., orders = orders, tz = tz),
    reasons = REASONS %>% as.character(),
    authority = AUTHORITY %>% as.character(),
    date_delisted = DATE.DELISTED %>% 
      parse_date_time(., orders = orders, tz = tz),
    comments = COMMENTS %>% as.character(),
    distribution = DISTRIBUTION %>% as.character(),
    fl_period = FL.PERIOD %>% as.character(),
    prev_taxonid = PREV.TAXONID %>% as.character(),
    prev_name = PREV.NAME %>% as.character(),
    fam_no = FAM.NO %>% as.character(),
    recovery_plan = R.PLAN %>% as.character(),
    recovery_plan_comments = RP.COMMENTS %>% as.character(),
    recovery_plan_expiry_date = RP.EXP.DATE %>% 
      parse_date_time(., orders = orders, tz = tz),
    active_ind = ACTIVE.IND %>% as.character(),
    deactivated_date = DEACTIVATED.DATE %>% 
      parse_date_time(., orders = orders, tz = tz),
    reason_deactivated = REASON.DEACTIVATED %>% as.character(),
    created_by = CREATED.BY %>% as.character(),
    created_date = CREATED.DATE %>% 
      parse_date_time(., orders = orders, tz = tz),
    modified_by = MODIFIED.BY %>% as.character(),
    modified_date = MODIFIED.DATE %>% 
      parse_date_time(., orders = orders, tz = tz),
    file_last_updated = FILE.LAST.UPDATED %>% 
      parse_date_time(., orders = orders, tz = tz),
    file_comments = FILE.COMMENTS %>% as.character(),
    file_no = FILE.NO %>% as.character()
  ) %>% 
  tibble::rowid_to_column("source_id") %>% 
  dplyr::arrange(name_id)

tfl_cons_listing[1:100, ] %>% DT::datatable(.)
readr::write_csv(tfl_cons_listing, path = here::here("data","tfl_cons_listing.csv"))

flcl_file <- here::here("data", "tfl_cons_listings.Rdata")
save(tfl_cons_listing, file=flcl_file)
if (file.exists(flcl_file)) load(flcl_file)

# ETL condensed criteria mixed with free text, therefore must be manually mapped
all_crit_map <- tfl_cons_listing %>%
  dplyr::group_by(list_code_wa) %>% 
  dplyr::distinct(criteria_code_wa) %>% 
  dplyr::arrange(list_code_wa, criteria_code_wa) %>% 
  dplyr::select(list_code_wa, criteria_code_wa)
readr::write_csv(all_crit_map, path = here::here("data","all_tfl_conscrit.csv"))
```


# Extract and transform conservation criteria Data from TSC
Conservation criteria Data from TSC is extracted through the TSC API and transformed into a useable
format for manual mapping to TFA.
``` {r extract_tsc_cons_criteria}
tsc_cons_list <- "conservationlist" %>% 
  wastdr::wastd_GET(api_url = prod) %>% 
  wastdr::wastd_parse()

tsc_cons_cat <- "conservationcategory" %>% 
  wastdr::wastd_GET(api_url = prod) %>% 
  wastdr::wastd_parse() %>% 
  dplyr::arrange(conservation_list, rank) %>% 
  dplyr::left_join(tsc_cons_list, by = c("conservation_list" = "id"))

tsc_cons_crit <- "conservationcriterion" %>% 
  wastdr::wastd_GET(api_url = prod) %>% 
  wastdr::wastd_parse() %>% 
  dplyr::arrange(conservation_list, rank) %>% 
  dplyr::left_join(tsc_cons_list, by = c("conservation_list" = "id"))
 
readr::write_csv(tsc_cons_crit,path = here::here("data","tsc_conscrit.csv"))

# Load manual mapping of TFA cons criteria from CSV file,  join to tfa_cons_crit
tfl_conscrit <- here::here("tfl_conscrit_w_tsc.csv") %>%
  readr::read_csv(
    col_types = cols(
      list_code_wa = col_character(),
      tsc_criteria = col_character(),
      criteria_code_wa = col_character(),
      assigned_as = col_character(),
      comments = col_character()
    )
  ) %>%
  dplyr::mutate(
    tsc_criteria = tsc_criteria %>% chr2int()
  )

glimpse(tfl_conscrit)

tfl_cons_listing_w_tsc<-tfl_cons_listing %>% 
  left_join(tfl_conscrit,by=c("list_code_wa","criteria_code_wa"))
```

# Load legacy data from TFA into TSC
In this section, the extracted and transformed data from TFA are loaded into TSC
using the TSC API. The code is collapsed; expand the code blocks to view the 
process.

## Taxonomy
Taxonomic names are uploaded to TSC directly from WACensus in a separate workbook.
The QA part of this workbook will compare TFA names to TSC names.

## Conservation Listings
This section loads conservation listings from TFA into TSC.

Conservation listings are entered in triplicate:

Level 1: Legislation

* WAWCA is the WA Wildlife Conservation Act, the legistation under which taxa 
  can be gazetted as specially protected.
* WAWCA has been superseded by the Biodiversity Conservation Act 2016 on 01/01/2019.
* All conservation listings were/are gazetted under the respective act (WAWCA or BCA).

Level 2: Gazette

* SPFN = "Wildlife Conservation (Specially Protected Fauna) Notice, Schedules"
* SPFN are the gazetted schedules.

Level 3: Assigned conservation categories and criteria

* DBCA has assigned actual conservation cat&crit to taxa following the IUCN definitions, 
  or similar cat&crit of own design before IUCN was published.
* These "IUCN" cons listings are the actual listings we want to preserve.

In future, only IUCN 2012 cat&crit will be assigned to taxa.
The list of all listed taxa will be published by the government gazette annually, 
which will set the life cycle status of our IUN listings to "gazetted".
The underlying legislation (from now on, the BCA) is evident out of the context 
and does not have to be entered as conservation listing into TSC.

ETL decisions for Fauna cons listings:

* Fauna WAPF = TSC WAPS = Priority species.
* Fauna RedList = TSC IUCN 2012 (or whichever earlier version was applicable 
  at the time).
* Fauna ActionPlan was decided to be discarded.
* Fauna has three empty listings (no list, no cat, no crit) - discard?

```{r tfl_cons_listing_upload, eval=F}
tsc_conscat_sp <- "conservationcategory" %>% 
  wastdr::wastd_GET(api_url = prod) %>% 
  magrittr::extract2("features") %>% 
  {tibble::tibble(
    tsc_category_id = purrr::map_int(., "id"),
    category_code = purrr::map_chr(., "code"),
    tsc_label = purrr::map_chr(., "label"),
    conservation_list = purrr::map_chr(., "conservation_list")
  )} %>% 
  dplyr::filter(conservation_list %in% c(1,2,3,4,5,6,7))

# List codes: "Priority"
# https://tsc.dbca.wa.gov.au/admin/conservation/conservationlist/6/
tfl_listing_cat_priority <- tfl_cons_listing_w_tsc %>% 
  dplyr::filter(list_code=="Priority") %>%
  dplyr::mutate(
    category_code=glue::glue("P{category_code}") %>% as.character) %>% 
  dplyr::left_join(
    dplyr::filter(tsc_conscat_sp, conservation_list == 6), 
    by="category_code") %>% 
  dplyr::transmute(
        source = 2, # TaxonGaz.source TFA 1, TFL 2, TEC 3
        source_id = source_id %>% as.character(),
        scope = 0, # state 0, cwth 1, intl 2, ap 3
        status = ifelse(is.na(date_delisted), 80, 90),
        effective_from = date_listed,
        effective_to = date_delisted,
        review_due = recovery_plan_expiry_date,
        taxon = name_id %>% as.numeric,
        category = tsc_category_id %>% as.list(),
        criteria = tsc_criteria
    ) %>% 
  dplyr::filter(!is.na(category))

tfl_listing_cat_priority %>% 
  wastdr::wastd_POST("taxon-conservationlisting", 
                     api_url = prod, verbose=TRUE)

# Active Priority listings:
tfl_listing_cat_priority %>% dplyr::filter(status==80) %>% nrow

# Inactive Priority listings:
tfl_listing_cat_priority %>% dplyr::filter(status==90) %>% nrow

# tfl_cl_priority[3888,] 
# taxonid 315936 does not exist, 31593 does - typo confirmed

# List codes: "EPBC" 
# https://tsc.dbca.wa.gov.au/admin/conservation/conservationlist/7/
tfl_listing_cat_epbc <- tfl_cons_listing_w_tsc %>% 
  dplyr::filter(list_code=="EPBC") %>% 
  dplyr::left_join(
    dplyr::filter(tsc_conscat_sp, conservation_list == 7), 
    by="category_code") %>% 
  dplyr::transmute(
        source = 2, # TaxonGaz.source TFA 1, TFL 2, TEC 3
        source_id = source_id %>% as.character(),
        scope = 1, # state 0, cwth 1, intl 2, ap 3
        status = ifelse(is.na(date_delisted), 80, 90),
        effective_from = date_listed,
        effective_to = date_delisted,
        review_due = recovery_plan_expiry_date,
        taxon = name_id %>% as.numeric,
        category = tsc_category_id %>% as.list(),
        criteria = tsc_criteria
    ) %>% 
  dplyr::filter(!is.na(category))

tfl_listing_cat_epbc %>% 
  wastdr::wastd_POST("taxon-conservationlisting", 
                     api_url = prod, verbose = TRUE)


# List codes: "WCA_1991" = IUCN 1994/2001/2012
# IUCN 1994 
# https://tsc.dbca.wa.gov.au/admin/conservation/conservationlist/5/change/
tfl_listing_cat_iucn1994 <- tfl_cons_listing_w_tsc %>% 
  dplyr::filter(list_code_wa=="IUCN_1994") %>% 
  dplyr::mutate(category_code=category_code_wa) %>% 
  dplyr::left_join(
    dplyr::filter(tsc_conscat_sp, conservation_list == 5), 
    by="category_code") %>% 
  dplyr::transmute(
        source = 2, # TaxonGaz.source TFA 1, TFL 2, TEC 3
        source_id = source_id %>% as.character(),
        scope = 0, # state 0, cwth 1, intl 2, ap 3
        status = ifelse(is.na(date_delisted), 80, 90),
        effective_from = date_listed,
        effective_to = date_delisted,
        review_due = recovery_plan_expiry_date,
        taxon = name_id %>% as.numeric,
        category = tsc_category_id %>% as.list(),
        criteria = tsc_criteria
    ) %>% 
  dplyr::filter(!is.na(category))

tfl_listing_cat_iucn1994 %>% 
  wastdr::wastd_POST("taxon-conservationlisting", 
                     api_url = prod, verbose = TRUE)

# IUCN 2001
# https://tsc.dbca.wa.gov.au/admin/conservation/conservationlist/4/change/
tfl_listing_cat_iucn2001 <- tfl_cons_listing_w_tsc %>% 
  dplyr::filter(list_code_wa=="IUCN_2001") %>% 
  dplyr::mutate(category_code=category_code_wa) %>% 
  dplyr::left_join(
    dplyr::filter(tsc_conscat_sp, conservation_list == 4), 
    by="category_code") %>% 
  dplyr::transmute(
        source = 2, # TaxonGaz.source TFA 1, TFL 2, TEC 3
        source_id = source_id %>% as.character(),
        scope = 0, # state 0, cwth 1, intl 2, ap 3
        status = ifelse(is.na(date_delisted), 80, 90),
        effective_from = date_listed,
        effective_to = date_delisted,
        review_due = recovery_plan_expiry_date,
        taxon = name_id %>% as.numeric,
        category = tsc_category_id %>% as.list(),
        criteria = tsc_criteria
    ) %>% 
  dplyr::filter(!is.na(category))

tfl_listing_cat_iucn2001 %>% 
  wastdr::wastd_POST("taxon-conservationlisting", 
                     api_url = prod, verbose=TRUE)


# IUCN 2012
# https://tsc.dbca.wa.gov.au/admin/conservation/conservationlist/3/change/
tfl_listing_cat_iucn2012 <- tfl_cons_listing_w_tsc %>% 
  dplyr::filter(list_code_wa=="IUCN_2012") %>% 
  dplyr::mutate(category_code=category_code_wa) %>% 
  dplyr::left_join(
    dplyr::filter(tsc_conscat_sp, conservation_list == 3), 
    by="category_code") %>% 
  dplyr::transmute(
        source = 2, # TaxonGaz.source TFA 1, TFL 2, TEC 3
        source_id = source_id %>% as.character(),
        scope = 0, # state 0, cwth 1, intl 2, ap 3
        status = ifelse(is.na(date_delisted), 80, 90),
        effective_from = date_listed,
        effective_to = date_delisted,
        review_due = recovery_plan_expiry_date,
        taxon = name_id %>% as.numeric,
        category = tsc_category_id %>% as.list(),
        criteria = tsc_criteria
    ) %>% 
  dplyr::filter(!is.na(category))

# There are none, skip upload
# tfl_cl_iucn2012 %>% 
#   wastdr::wastd_POST("taxon-conservationlisting", api_url = prod)
```

## Occurrences

## Conservation Documents

## Fire History

## Conservation Threats and Actions


# Extract and transform migrated Data from TSC
Data from TSC is extracted through the TSC API and transformed into a useable
format.

Expand the code block below to view the process.

```{r load_tsc, message=FALSE}
# TSC contains taxonomic names from WACensus
tsc_taxa_raw <- "taxon" %>%
  wastdr::wastd_GET(api_url=prod, query = list(paraphyletic_groups__icontains=21)) 

tsc_taxa <- tsc_taxa_raw %>%
  wastdr::wastd_parse() %>% 
  dplyr::mutate(name_id = name_id %>% as.character) 

# tsc_comm <- "community" %>% 
#   wastdr::wastd_GET(api_url=prod) %>% 
#   wastdr::wastd_parse() # needs an upgrade to handle complex geometries

tsc_tcl_raw <- wastdr::wastd_GET("taxon-conservationlisting") 

tsc_tcl <- tsc_tcl_raw %>% wastdr::parse_taxon_conservationlisting() %>% 
  dplyr::rename(name_id = taxon) %>% 
  dplyr::mutate(name_id = name_id %>% as.character) 

tsc_data <- tsc_tcl %>% 
  dplyr::inner_join(tsc_taxa,by = "name_id")
```

```{r save_point, message=FALSE}
if (file.exists("tracks.Rda")){
    load("tracks.Rda")
} else {
  save(
    tsc_taxa,
    # tsc_comm
    # TODO add all data objects to keep in here
    file = here::here("data", "fauna.Rda")
  )
}
```


