---
title: "Data ETL: Fauna"
author: "Florian Mayer and Milly Piggott, DBCA"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
  pdf_document: default
---

# Context
QA TFA fauna dataset and TSC fauna dataset and make comparisons to check all data has been transfered correctly to TSC.

[TFL Fauna dataset](https://data.dpaw.wa.gov.au/dataset/threatened-and-priority-fauna-database) 
on the [departmental data catalogue](https://data.dpaw.wa.gov.au/).

[TSC Fauna database](https://tsc.dbca.wa.gov.au/species/?paraphyletic_groups=20&is_terminal_taxon=true) 
on the [TSC catalogue](https://tsc.dbca.wa.gov.au/).

The code for all SCB workbooks is under version control at [github](https://github.com/dbca-wa/scarab-scripts).

# Data
The zip archive containing the MS Access database is downloaded from the data 
catalogue, the archive is extracted, and a database connection is opened.

```{r fauna_data_load, message=FALSE}
fauna_data <- dl_mdbzip("66efb68d-8f05-4bfc-af14-5d1a381d0cf2")
```

# Lookup tables
Lookup table columns are renamed to facilitate joining.

```{r lookup_tables}
# Lookup tables
shires <- fauna_data$Shires %>%
  dplyr::transmute(
    lga_id = LGACode %>% as.integer(),
    shire_name = LGAName %>% as.character(),
    shire_label = LGAName2 %>% as.character()
  )

tenure <- fauna_data$`Land Tenure` %>%
  dplyr::transmute(
    tenure_id = TenCode %>% as.character(),
    tenure = Tenure %>% as.character()
  )

vegetation_types <- fauna_data$`Vegetation Types` %>%
  dplyr::mutate(
    vegetation_type_id = VegCode %>% as.character(),
    vegetation_type = VegTypeName %>% as.character()
  )

certainty <- fauna_data$Certainty %>%
  dplyr::transmute(
    certainty_id = Code %>% as.integer(),
    certainty = Certainty %>% as.character()
  )

users <- tibble::tibble(
  username = c(
    "AmyM", "AbbyT", "BrianaWingfield",
    "ChristineFreegard", "GeorginaA", "KellieMantle",
    "NickyMarlow", "PeterMawson", "PeterOrell"
  ),
  initials = c(
    "AM", "AT", "BW",
    "CF", "GA", "KM",
    "NM", "PM", "PO"
  ),
  name = c(
    "Amy Mutton", "Abby Thomas", "Briana Wingfield",
    "Christine Freegard", "Georgina Anderson", "Kellie Mantle",
    "Nicky Marlow", "Peter Mawson", "Peter Orell"
  )
)

phylo_class <- fauna_data$`Class List` %>%
  dplyr::transmute(
    phylo_class_id = ClassCode %>% as.character(),
    phylo_class_common_name = ClassName %>% as.character(),
    phylo_class_scientific_name = Class %>% as.character(),
    order = ClassOrder %>% as.character()
  ) %>%
  dplyr::arrange(order)

phylo_group <- fauna_data$`Group List` %>%
  dplyr::transmute(
    phylo_group_id = Group %>% as.integer(),
    phylo_group_name = GroupName %>% as.character(),
    phylo_class_id = Class %>% as.character()
    # order = GroupOrder %>% as.character()
  ) %>%
  dplyr::left_join(phylo_class, by = "phylo_class_id")
# skimr::skim(group)

species_notes <- fauna_data$`Species Notes` %>%
  dplyr::transmute(
    species_code = SpCode %>% as.character(),
    species_notes = Notes %>% as.character()
  )

resolution <- fauna_data$Resolution %>%
  dplyr::transmute(
    resolution_id = ResCode %>% as.integer(),
    resolution_geographic_coords = ResolutionLL %>% as.character(),
    resolution_projected_coords = ResolutionUTM %>% as.character()
  )

resolution_m <- tibble::tibble(
    resolution_id = c(1,2,3,4,5,6),
    accuracy = c(50,100,1000, 10000, 50000, 100000)
)

category <- fauna_data$`Category List` %>%
  dplyr::transmute(
    category_id = Category %>% as.character(),
    phylogenetic_category = CategoryName %>% as.character()
  )
```
