---
title: "Data ETL from ODK Central: Flora Quadrats"
author: "Florian Mayer, DBCA"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
  pdf_document: default
---


# Background
This workbook demonstrates how to access data from ODK Central through the
OData service endpoint.

As an example, we show three forms from Flora Quadrat Surveys.

Data is uploaded to the 
[ODK Central sandbox](https://sandbox.central.opendatakit.org/#/projects/14/).
User and app accounts can be created on request.

This example requires the email address (=ODK Central username) and password 
to be set as R env vars `ODKC_UN` and `ODKC_PW` for the ODK Central server.
A good place for those is `~/.Rprofile`.

```{r setup, message=F}
if (file.exists("~/.Rprofile")) source("~/.Rprofile")
# library(OData)  # didn't work, went with httr
# remotes::install_github("r-lib/rlang")
# remotes::install_github("tidyverse/tidyr")
library(ckanr)
# library(Hmisc)  # MS Access
library(glue)
library(httr)
library(tidyr)
library(dplyr)
library(magrittr)
library(tibble)
library(purrr)
library(leaflet)
library(wastdr)
# Our example OData endpoint, plus /Submissions
base_url <- "https://sandbox.central.opendatakit.org/v1/projects/14/forms/"
form_id <- "build_Flora-Quadrat-0-1_1558330379"
data_url <- glue::glue("{base_url}{form_id}.svc")
```


# Data ETL

## Extract: accessing the data
In this section, form submissions are downloaded from ODK Central and transformed
into a rectangular shape.

```{r helpers}
#' Retrieve metadata from an OData URL ending in .svc as list of lists
#'
#' @param url The OData URL, ending in .svc, no trailing slash
#' @param un The ODK Central username (an email address), 
#'           default: Sys.getenv("ODKC_UN").
#'           Add to your ~/.Rprofile: Sys.setenv(ODKC_UN="...@...")
#' @param pw The ODK Central password,, 
#'           default: Sys.getenv("ODKC_PW").
#'           Add to your ~/.Rprofile: Sys.setenv(ODKC_PW="...")
#' @returns A list of two named lists, DataServices and .attrs (Version).
#'          DataServices contains the dataset definition.
get_metadata <- function(
  url, 
  un=Sys.getenv("ODKC_UN"),
  pw=Sys.getenv("ODKC_PW")
){
  glue::glue("{url}/$metadata") %>% 
  httr::GET(
    httr::add_headers(Accept = "application/xml"), 
    httr::authenticate(un, pw)
  ) %>% 
  httr::content(.) %>% 
  XML::xmlParse(.) %>% 
  XML::xmlToList(.) 
}

#' Retrieve /Submissions from an OData URL ending in .svc as list of lists
#' 
#' @param url The OData URL, ending in .svc, no trailing slash
#' @param un The ODK Central username (an email address), 
#'           default: Sys.getenv("ODKC_UN").
#'           Add to your ~/.Rprofile: Sys.setenv(ODKC_UN="...@...")
#' @param pw The ODK Central password,, 
#'           default: Sys.getenv("ODKC_PW").
#'           Add to your ~/.Rprofile: Sys.setenv(ODKC_PW="...")
#' @importFrom glue glue
#' @importFrom httr add_headers authenticate content
#' @returns A list of two named lists, value and context.
#           $value contains the submissions, which can be "rectangled"
#'          using `tidyr::unnest_wider("element_name")`.
get_submissions <- function(
  url, 
  un=Sys.getenv("ODKC_UN"),
  pw=Sys.getenv("ODKC_PW")
){
  glue::glue("{url}/Submissions") %>% 
  httr::GET(
    httr::add_headers(Accept = "application/json"), 
    httr::authenticate(un, pw)
  ) %>% 
  httr::content(.)
}

#' Download attachments and return the local path
#' 
#' @param base_url The ODK Central OData base url,
#'                 e.g. "https://sandbox.central.opendatakit.org/v1/projects/14/forms/"
#' @param form_id The ODK form ID, e.g. "build_Flora-Quadrat-0-1_1558330379"
#' @param submission_uuid The ODK submission UUID, an MD5 hash
#' @param attachment_filename The ODK form attachment filename, 
#'                            e.g. "1558330537199.jpg"
#' @param local_dir The local folder to save the downloaded files to, 
#'                  default: "attachments"
#' @param un The ODK Central username (an email address), 
#'           default: Sys.getenv("ODKC_UN").
#'           Add to your ~/.Rprofile: Sys.setenv(ODKC_UN="...@...")
#' @param pw The ODK Central password,, 
#'           default: Sys.getenv("ODKC_PW").
#'           Add to your ~/.Rprofile: Sys.setenv(ODKC_PW="...")
dl_attachment <- function(
  base_url,
  form_id,
  submission_uuid, 
  attachment_filename,
  local_dir="attachments",
  un=Sys.getenv("ODKC_UN"),
  pw=Sys.getenv("ODKC_PW"),
  overwrite=FALSE
  ){
  
  # Create local destination dir attachments/uuid
  dest_dir <- glue::glue("{local_dir}/{submission_uuid}")
  dir.create(dest_dir, recursive = T)
  
  source_url <- glue::glue("{base_url}{form_id}/",
               "submissions/{submission_uuid}/",
               "attachments/{attachment_filename}")
  
  dest_path <- glue::glue("{dest_dir}/{attachment_filename}")
  
  # Keep existing
  if (!file.exists(dest_path) && !is.na(attachment_filename)){
    httr::GET(
      source_url, 
      httr::authenticate(un, pw), 
      write_disk(dest_path, overwrite=T)
    )
  }
  
  # Return the relative path to the attechment
  return(dest_path %>% as.character)
}
```


```{r extract, message=F, warning=F}
quadrat_metadata <- get_metadata(data_url)
d <- get_submissions(data_url)
data <- tibble::tibble(value=d$value) %>% 
  tidyr::unnest_wider(value) %>% 
  dplyr::rename(uuid=`__id`) %>% 
  tidyr::unnest_wider(`__system`) %>% 
  tidyr::unnest_wider(meta) %>% 
  tidyr::unnest_wider(location) %>% 
  tidyr::unnest_wider(corner1) %>% 
  tidyr::unnest_wider(coordinates) %>% 
  dplyr::rename(longitude=`...1`, latitude=`...2`, altitude=`...3`) %>% 
  dplyr::select(-"type") %>% 
  tidyr::unnest_wider(habitat) %>%   
  tidyr::unnest_wider(vegetation_structure) %>%   
  tidyr::unnest_wider(perimeter) %>% 
  tidyr::unnest_wider(coerner2) %>% 
  tidyr::unnest_wider(coordinates) %>%
  dplyr::select(-"type") %>% 
  dplyr::rename(longitude_c2=`...1`, latitude_c2=`...2`, altitude_c2=`...3`) %>% 
  tidyr::unnest_wider(coerner3) %>% 
  tidyr::unnest_wider(coordinates) %>%
  dplyr::select(-"type") %>% 
  dplyr::rename(longitude_c3=`...1`, latitude_c3=`...2`, altitude_c3=`...3`) %>% 
  tidyr::unnest_wider(coerner4) %>% 
  tidyr::unnest_wider(coordinates) %>%
  dplyr::select(-"type") %>% 
  dplyr::rename(longitude_c4=`...1`, latitude_c4=`...2`, altitude_c4=`...3`) %>%
  dplyr::mutate(
    quadrat_photo_local = dl_attachment(base_url, form_id, uuid, quadrat_photo),
    morphological_type_photo_local = dl_attachment(base_url, form_id, uuid, morphological_type_photo),
    mudmap_photo_local = dl_attachment(base_url, form_id, uuid, mudmap_photo)
  )
```

## Analyse
```{r}
data %>% DT::datatable(.)
```

```{r}
leaflet::leaflet(width = 800, height = 600) %>%
  leaflet::addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
  leaflet::addProviderTiles("OpenStreetMap.Mapnik", group = "Place names") %>%
  leaflet::clearBounds() %>% 
  leaflet::addAwesomeMarkers(
    data = data,
    lng = ~longitude, lat = ~latitude,
    icon = leaflet::makeAwesomeIcon(text = "Q", markerColor = "red"),
    label = ~glue::glue('{area_name} {encounter_start_datetime}'),
    popup = ~glue::glue(
      "<h3>{area_name}</h3>",
      "Survey start {encounter_start_datetime}</br>",
      "Reporter {reporter}</br>",
      "Device {device_id}</br>",
      "<h5>Site</h5>",
      '<div><img src="{quadrat_photo_local}"',
      ' height="150px" alt="Quadrat photo"></img></div>',
      "<h5>Mudmap</h5>",
      '<div><img src="{mudmap_photo_local}',
      ' height="150px" alt="Mudmap"></img></div>',
      "<h5>Habitat</h5>",
      "Morphological type: {morphological_type}</br>",
      '<div><img src="{morphological_type_photo_local}"',
      'height="150px" alt="Morphological type"></img></div>',
      "Veg class: {vegclass_placeholder}</br>"
    ),
    clusterOptions = markerClusterOptions()
  ) %>%
  addLayersControl(
    baseGroups = c("Aerial", "Place names"),
    options = layersControlOptions(collapsed = FALSE)
  )
```
