---
title: "Let's get WAStD"
author: "Florian Mayer, DBCA"
date: "31/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(wastdr)
datafile <- here::here("data/tae.RData")
```

# Setup wastdr
Read how to [setup](https://dbca-wa.github.io/wastdr/articles/setup.html) wastdr
to the TSC API.

# Download Taxon Occurrences from TSC
We get Taxon ocurrences with their point representations from TSC through the 
API endpoint [occ-taxon-points](https://tsc.dbca.wa.gov.au/api/1/occ-taxon-points/).

Currently, all Fauna and Flora occurrences are georeferenced only with point 
coordinates, not polygons.

We then get TSC Areas and split them into DBCA Regions and Districts.
Caveat: Regions and Districts to not comprehensively cover WA.

Lastly, we get a list of taxonomic names from TSC to resolve taxon IDs from
taxon occurrences to names.

Occurences are presented as `occ` with DBCA Region and District names (where
applicable), plus taxonomic names.

```{r download, eval=F}
tsc_tae <- wastdr::wastd_GET("occ-taxon-points") # 137k records

tsc_taxa <- wastdr::wastd_GET("taxon") %>%
    wastdr::wastd_parse() %>%
    dplyr::mutate(name_id = name_id %>% as.character(), taxon = pk)

tsc_areas <- wastdr::wastd_GET("area") %>%
      magrittr::extract2("features") %>%
      geojsonio::as.json() %>%
      geojsonsf::geojson_sf()

save(tsc_tae, tsc_taxa, tsc_areas, file=datafile, compress="xz")
```

```{r munge}
if (fs::file_exists(datafile)) load(datafile)

regions <- tsc_areas %>%
  dplyr::filter(area_type == "Region") %>%
  dplyr::transmute(region_id = pk, region_name = name)

districts <- tsc_areas %>%
  dplyr::filter(area_type == "District") %>%
  dplyr::transmute(district_id = pk, district_name = name)

occ <- tsc_tae %>%
  magrittr::extract2("features") %>%
  geojsonio::as.json() %>%
  geojsonsf::geojson_sf() %>% 
  sf::st_join(regions) %>% 
  sf::st_join(districts) %>% 
  dplyr::left_join(tsc_taxa, by="taxon")

# save(occ, file=here::here("data/occ.RData"), compress="xz")
# load(here::here("data/occ.RData"))
```

# Visualise occurrences

## Species occurrences by DBCA Region

```{r occ_by_reg}
occ_by_region <- occ %>% 
    dplyr::group_by(region_name, taxonomic_name) %>% 
    dplyr::tally()  %>% 
    wastdr::sf_as_tbl() %>% 
    tidyr::pivot_wider(
        id_cols=taxonomic_name,
        names_from = region_name,
        values_from = n,
        values_fill = NA
    )

occ_by_region %>% reactable::reactable(sortable = T, filterable = T)
```

## Species occurrences by DBCA District

```{r occ_by_dist}
occ_by_district <- occ %>% 
    dplyr::group_by(district_name, taxonomic_name) %>% 
    dplyr::tally()  %>% 
    wastdr::sf_as_tbl() %>% 
    tidyr::pivot_wider(
        id_cols=taxonomic_name,
        names_from = district_name,
        values_from = n,
        values_fill = NA
    )

occ_by_district %>% reactable::reactable(sortable = T, filterable = T)
```

