---
title: "Occurrence analysis"
author: "Florian Mayer, DBCA"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
# devtools::install_github("ropenscilabs/skimr")
# devtools::install_github("tidyverse/readr")
# devtools::install_github("brianb/mdbtools")
# devtools::install_github("ropensci/ckanr")
# devtools::install_github('thomasp85/gganimate')
library(tidyverse)
library(jsonlite)
# library(geojsonio)
library(ckanr)

library(magrittr)
library(lubridate)
library(skimr)
library(grDevices)

library(leaflet)
library(sf)
library(geojsonsf)
library(mapview)
library(maps)
library(gganimate)
library(ggthemes)
library(mapproj)
library(RColorBrewer)
# library(gapminder)

# Configure ckanr to data.dpaw.wa.gov.au
ckanr::ckanr_setup(url = Sys.getenv("CKAN_URL"), key = Sys.getenv("CKAN_API_KEY"))
source("helpers.R")
dev <- "http://localhost:8220/api/1/"
uat <- "https://tsc-uat.dbca.wa.gov.au/api/1/"
prod <- "https://tsc.dbca.wa.gov.au/api/1/"
```

# Context
This workbook analyses occurrence data of threatened species and communities as
extracted from the legacy databases.

Currently, data snapshots from the data catalogue as exported by the ETL workbooks are used. 
Soon, we will use occurrence data from BioSys.

# Usage
This workbook serves two purposes.

The data analyst (Florian) runs this workbook to generate EOO (extent of occurrence)
polygons and upload them to [TSC](https://tsc.dbca.wa.gov.au/). 

At the same time, this workbook aims to document and explain the process, 
so that TSC stakeholders can verify that this process is correct and produces the desired outputs.

The data curators will contribute feedback for the analyses and visualisations in this workbook
to be improved and expanded, which the analyst will implement.

While a rendered copy of this workbook (for stakeholders to read) lives on the DBCA data catalogue 
([dataset link](https://data.dpaw.wa.gov.au/dataset/threatened-species-and-communities-occurrence-analysis) and
[fullscreen link](https://data.dpaw.wa.gov.au/dataset/a9c0fc22-3fd0-4896-9aba-95ad16e0c02a/resource/56335544-72cf-4ebd-bb2c-bc246ec96080/download/occ_etl.html)), its source (for analysts to run) is under version control [here](https://github.com/dbca-wa/scarab-scripts).

# Data download
* Now: Occurrence data snapshots from data catalogue
* Soon: Occurrence data from Biosys

```{r load_data, message=F}
fauna_occ_rid <- "b73ae0de-f243-4bb3-bbce-25b65bb98e66"
flora_occ_rid <- "b2e7f5a9-8ffb-41c4-bd11-ab090ccb71ef"
comm_occ_rid <- ""
fauna_occ <- ckanr::resource_show(fauna_occ_rid)$url %>% readr::read_csv()
flora_occ <- ckanr::resource_show(flora_occ_rid)$url %>% readr::read_csv()
```

# A first look

```{r tally_occ}
fauna_tally <- fauna_occ %>%
  dplyr::group_by(name_id, scientific_name) %>%
  dplyr::tally() %>%
  dplyr::arrange(desc(n))

fauna_tally %>% head(n = 20) %>% kable(.)

flora_tally <- flora_occ %>%
  dplyr::group_by(name_id, species_name) %>%
  dplyr::tally() %>%
  dplyr::arrange(desc(n))

flora_tally %>% head(n = 20) %>% kable(.)
```

There are `r nrow(fauna_tally)` fauna taxa and 
`r nrow(flora_tally)` flora taxa with recorded occurrences.

# Helper functions
A few helper functions to extract convex hulls from given points.

```{r helpers}
#' Return the convex hull plus minimal buffer around all given points.
#'
#' A minimal buffer is included to turn single point observations
#' into actual polygons.
#' @param lonlat_cols A vector of lon and lat column names,
#'   default: c("longitude", "latitude")
#' @param crs The coordinate reference system ID (4283=GDA94)
#' @return An object of class "sfc_POLYGON" "sfc" (package sf)
lonlat_to_convex_hull <- function(data,
                                  lonlat_cols=c("longitude", "latitude"),
                                  crs=4283) {
  data %>%
    sf::st_as_sf(coords = lonlat_cols, crs = crs, agr = "constant") %>%
    sf::st_union(.) %>%
    sf::st_convex_hull(.) %>%
    sf::st_buffer(., 0.00001)
}

#' Convert occurrence records into a dataframe of distinct name_id:eoo
#'
#' **Data** must have the columns `name_id`, `latitude`, and `longitude`.
#' **Filter** out occurrences of invalid NameIDs (NA and 0).
#' **Nest** occurrence data by name_id to generate two columns,
#'   `name_id` and `data` (all records of that `name_id`).
#' **Map** `lonlat_to_convex_hull` over data column to generate convex
#'   hulls (class `sfc`) from all points in `data` column.
#' **Map** `sfc_geojson` to convert the convex hulls into GeoJSON.
#' @param data A dataframe with at least the columns
#'   `name_id`, `latitude`, and `longitude`.
#' @return A dataframe with columns `name_id`, `data`, `eeo_sfc`, and `eeo`:
#'      name_id: one row per distinct name_id in input dataframe.
#'      data: the nested data of all rows with that name_id.
#'      eeo_sfc: the extent of occurrence as object of class `sfc`.
#'      eeo: the extent of occurrence as GeoJSON string.
#' @examples
#'   \notrun{
#'   data %>% make_eoo %>% mapview(.$eoo_sfc)
#'   data %>% make_eoo %>% dplyr::select(-data, -eoo_sfc) %>% wastdr::wastd_POST("taxon")
#'   }
make_eoo <- function(data) {
  data %>%
    dplyr::filter(
      !name_id %in% c(NA, 0),
      !is.na(longitude),
      !is.na(latitude)
    ) %>%
    tidyr::nest(-name_id) %>%
    dplyr::mutate(
      eoo_sfc = purrr::map(data, lonlat_to_convex_hull),
      eoo = purrr::map(eoo_sfc, geojsonsf::sfc_geojson)
    )
}

#' Return the convex hull around all points with given name_id
#'
#' @param data A tibble or data.frame with columns
#'   `longitude`, `latitude`, `name_id`
#' @param nid A numeric name_id, default: null (use all rows)
#' @param nid_col The column name of the NameID column, default: "name_id"
#' @param lonlat_cols A vector of lon and lat column names,
#'   default: c("longitude", "latitude")
#' @param crs The coordinate reference system ID (4283=GDA94)
#' @return An object of class "sfc_POLYGON" "sfc" (package sf)
#'   if points are given, else NULL.
eoo_polygon <- function(data,
                        nid=NULL,
                        nid_col="name_id",
                        lonlat_cols=c("longitude", "latitude"),
                        crs=4283) {
  if (is.null(nid)) {
    selected <- data
  } else {
    selected <- dplyr::filter(data, (!! rlang::sym(nid_col)) == nid)
  }
  if (nrow(selected) == 0) return(NULL)
  lonlat_to_convex_hull(selected, lonlat_cols = lonlat_cols, crs = crs)
}

anim_fauna <- function(data, title, nid){
  d <- data %>% 
    dplyr::mutate(year = lubridate::year(datetime)) %>% 
    dplyr::filter(name_id == nid)
    
  ggplot() +
  borders("world", colour = "gray75", fill = "gray95") +
  coord_equal(xlim = c(110, 130), ylim = c(-40,-12)) +
  ggthemes::theme_map(base_size = 14) + 
  ggtitle(title, subtitle = "{lubridate::year(frame_time)}") +
  geom_point(data = d,
            aes(x = longitude, 
                y = latitude, 
                size = number_seen, 
                colour = observation_type), 
            alpha = 0.7,
            show.legend = FALSE) +
  scale_size(range = c(2,8)) +
  transition_time(o_date) +
  ease_aes('linear')
  
  # anim_save(glue::glue("data/occ_{nid}.gif"))
}
```

# Examples

## All occurrences
Excluding flora data with missing coordinates.

```{r ex_all_occ}
fauna_eoo_all <- eoo_polygon(fauna_occ)
mapview(fauna_eoo_all)

flora_eoo_all <- flora_occ %>%
  dplyr::filter(!is.na(longitude), !is.na(latitude)) %>%
  eoo_polygon()
mapview(flora_eoo_all)
```

The next sections show a handful of example taxa and their calculated EOOs.

Once this workbook has generated and uploaded the EOO polygons to TSC, 
the EOOs shown in the following examples should be the same as the EOO shown 
in TSC (see the respective TSC GUI links to view them in TSC).

## Leipoa ocellata
Malleefowl

* TSC GUI: https://tsc.dbca.wa.gov.au/species/24557/
* TSC API: https://tsc.dbca.wa.gov.au/api/1/taxon/?name_id=24557

```{r ex_malleefowl, warning=F}
eoo_24557 <- eoo_polygon(fauna_occ, nid = 24557)

# see https://r-spatial.github.io/sf/articles/sf5.html
# ggplot() + geom_sf(data = eoo_24557)
mapview::mapview(eoo_24557)

anim_fauna(fauna_occ, "Malleefowl", 24557)
```

## Calyptorhynchus latirostris
Carnaby's Black Cockatoo: 14k occurrences.

* TSC GUI: https://tsc.dbca.wa.gov.au/species/24734/
* TSC API: https://tsc.dbca.wa.gov.au/api/1/taxon/?name_id=24734

```{r ex_carnabys, warning=F}
eoo_24734 <- eoo_polygon(fauna_occ, nid = 24734)
mapview::mapview(eoo_24734)
anim_fauna(fauna_occ, "Carnaby's Black Cockatoo", 24734)


anim_fauna(fauna_occ, "Woylie", 24162)
```

## Banksia brownii

* TSC GUI: https://tsc.dbca.wa.gov.au/species/1806/
* TSC API: https://tsc.dbca.wa.gov.au/api/1/taxon/?name_id=1806

```{r ex_bbrownii}
eoo_1806 <- flora_occ %>% 
    dplyr::filter(!is.na(longitude), !is.na(latitude)) %>% 
    eoo_polygon(nid = 1806)
mapview::mapview(eoo_1806)
```

## Banksia goodii

* TSC GUI: https://tsc.dbca.wa.gov.au/species/1818/
* TSC API: https://tsc.dbca.wa.gov.au/api/1/taxon/?name_id=1818

```{r ex_bgoodii}
eoo_1818 <- flora_occ %>% 
    dplyr::filter(!is.na(longitude), !is.na(latitude)) %>% 
    eoo_polygon(nid = 1818)
mapview::mapview(eoo_1818)
```

## TSC upload example
Writing an `sf` object to a GeoJSON string:
https://cran.r-project.org/web/packages/geojsonsf/vignettes/geojson-sf-conversions.html

```{r ex_upload_one_eoo, eval=F}
one_taxon_eoo <- tibble::tibble(
  name_id = 24162,
  eoo = eoo_polygon(fauna_occ, nid = 24162) %>% sfc_geojson()
)
wastdr::wastd_POST(one_taxon_eoo, "taxon", verbose = T)
```

# Data QA
### Fauna occurrences missing a NameID

```{r}
fauna_occ_no_nameid <- fauna_occ %>% 
    filter(name_id == 0) %>% 
    select(scientific_name) %>% 
    unique()
fauna_occ_no_nameid %>% kable
```

There are `r nrow(fauna_occ_no_nameid)` distinct scientific fauna names without NameID.
Some are known to TSC: E.g. the Western Brush Wallaby, Notamacropus irma, 
NameID [48022](https://tsc.dbca.wa.gov.au/species/48022/).
These records need to be fixed in the original Threatened Fauna database.

```{r}
flora_occ_no_nameid <- flora_occ %>% 
    filter(name_id == 0) %>% 
    select(species_name) %>% 
    unique()
flora_occ_no_nameid %>% kable
```

There are `r nrow(flora_occ_no_nameid)` distinct scientific flora names without NameID.

## Missing coordinates

```{r}
fauna_occ_no_latlng <- fauna_occ %>% 
    filter(is.na(latitude) | is.na(longitude))
fauna_occ_no_latlng %>% head %>% kable
```

There are `r nrow(fauna_occ_no_latlng)` distinct fauna records with missing latitude or longitude.

```{r}
flora_occ_no_latlng <- flora_occ %>% 
    filter(is.na(latitude) | is.na(longitude))
flora_occ_no_latlng %>% head %>% kable
```

There are `r nrow(flora_occ_no_latlng)` distinct flora records with missing latitude or longitude.
These records need to be fixed in the original Threatened Flora database.

## NameID mismatch
TODO diagnose. Lookup scientific names in TSC, get name_id, compare to name_id here.

## Trust level of occurrence records
Currently, we assume that all occurrence records in the legacy databases are trustworthy.
If this weren't the case, this analysis should exclude the non-trustworthy records.

# EOO to TSC upload
This section generates and uploads the EOOs to TSC through its API using `wastdr`.
For performance reasons, we'll run the steps in batches on the entire 
fauna, flora (and soon: TEC) datasets.

**Note** Only occurrence records with valid NameIDs and coordinates are considered.

* **Data** must have the columns `name_id`, `latitude`, and `longitude`.
* **Filter** out occurrences of invalid NameIDs (NA and 0).
* **Nest** occurrence data by name_id to generate two columns, 
  `name_id` and `data` (all records of that `name_id`).
* **Map** `lonlat_to_convex_hull` over data column to generate convex 
  hulls (class `sfc`) from all points in `data` column.
* **Map** `sfc_geojson` to convert the convex hulls into GeoJSON.
* **Discard** columns `data` and `eeo_sfc`.
* **POST** to WAStD.

Crunch time.

```{r calculate_eoo}
fauna_eoo <- fauna_occ %>% make_eoo()
flora_eoo <- flora_occ %>% make_eoo()
```

Does that look about right?

```{r map_eoo}
mapview::mapview(head(fauna_eoo$eoo_sfc))
mapview::mapview(head(flora_eoo$eoo_sfc))

# Banksia brownii, NameID 1806
flora_eoo %>% 
    filter(name_id==1806) %>% 
    magrittr::extract2("eoo_sfc") %>% 
    mapview::mapview(.)

flora_eoo %>% 
    # dplyr::filter(name_id==1806) %>%
    dplyr::select(-data, -eoo_sfc) %>%
    wastdr::wastd_POST("taxon", verbose = T, api_url = dev)
# # Banksia brownii, NameID 1806
# flora_eoo %>% 
#     filter(name_id==1806) %>% 
#     magrittr::extract2("eoo_sfc") %>% 
#     mapview::mapview(.)
# 
# flora_eoo %>% 
#     dplyr::filter(name_id==1806) %>%
#     dplyr::select(-data, -eoo_sfc) %>%
#     wastdr::wastd_POST("taxon", verbose = T, api_url = uat)
```

Upload to TSC - run manually. This could overwrite data or time out (depending on network load).

```{r upload_eoo, eval=F}
fauna_eoo %>%
  dplyr::select(-data, -eoo_sfc) %>%
  wastdr::wastd_POST("taxon", verbose = T, api_url = prod)

flora_eoo %>%
  dplyr::select(-data, -eoo_sfc) %>%
  wastdr::wastd_POST("taxon", verbose = T, api_url = prod)
```

# TEC
* Boundaries of community occurrences are available as Shapefiles.
* QGIS: Open shapefile, Fix geometries, Minimum bounding geometry of type "convex hull", group by "COM_ID".
* Resulting layer: save as GeoJSON, upload to `data/tec_eoo.geojson` (relative to workbook).

Note: TSC's API only accepts GeoJSON strings for the field EEO, not (Geo)JSON objects.

Note: A custom QGIS model wraps the QGIS steps into one GUI dialogue.

```{r, eval=F}
tec <- geojsonsf::geojson_sf("data/tec_eoo.geojson")
tibble::tibble(code = tec$COM_ID, eoo = geojsonsf::sfc_geojson(tec$geometry)) %>% 
    wastdr::wastd_POST("community", verbose=T, api_url = uat)
```


# Occurrences to TSC
A single record, or a list of (one or) many records, can be uploaded to the TSC API as follows.

```{r eval=F}
# 25755
taxonpoint = rjson::fromJSON('[{"taxon": 25757, "code": "Condingup", "description": "Pine plantations off Fisheries Rd east of Condinup\nOpportunistic sighting\nDay sighting\nHeard/Call\nSaw a flock of approximately 200 birds in the pine plantation\nObserved by Abby Thomas, DBCA Technical Officer, Albany",    "source": 10,    "source_id": "94654",    "encountered_on": "2018-04-16T15:00:00+08:00",    "encountered_by": 1,    "area_type": 30,    "accuracy": 1000.0,    "point":"SRID=4326;POINT(119.732234 -32.35756)"}]')
wastdr::wastd_POST(taxonpoint, serializer = "occ-taxon-points", verbose = T, api_url = dev)
```

```{r fauna_occ_tsc}
fauna_tae <- fauna_occ %>%
  dplyr::filter(!is.na(name_id)) %>% 
  dplyr::transmute(
    taxon = name_id,
    code = location_name,
    description = glue::glue(
        "{observation_method} ({observation_type}) by {observer_name} ({observer_role}) {observer_address} {observer_phone}. ",
        "Certainty {certainty}. ",
        "Number seen: {number_seen}. ",
        "Site comments: {site_comment}. ",
        "Observation comments: {observation_comments}. ",
        "Curator comments: {comments}"),
    source = 10,
    source_id = id,
    encountered_on = glue::glue("{datetime}+08:00"),
    encountered_by = 1,
    area_type = 30,
    accuracy = accuracy,
    point = glue::glue("SRID=4326;POINT({longitude} {latitude})")
  )

# db transaction spans entire batch, data only visible in TSC on completed POST
fauna_tae10 = fauna_tae[20001:62901,]
wastdr::wastd_POST(fauna_tae10, serializer = "occ-taxon-points", api_url = dev)
```

```{r flora_occ_tsc}
flora_tae <- flora_occ %>%
  dplyr::filter(!is.na(name_id)) %>% 
  dplyr::filter(!is.na(observation_date)) %>%
  dplyr::filter(!is.na(latitude)) %>%
  dplyr::filter(!is.na(longitude)) %>%
  dplyr::transmute(
    taxon = name_id,
    code = pop_id,
    name = glue::glue("{pop_id}-{subpop_code}"),
    description = glue::glue(
        "Population {pop_id}, Subpopulation {subpop_code}. ",
        "Location: {location}. Nearest known place: {nearplace}. "
        ),
    source = 11,
    source_id = sheetno,
    encountered_on = glue::glue("{observation_date} 00:00:00+08:00"),
    encountered_by = 1,
    area_type = 20, # pop:20, subpop: 21
    accuracy = 1000,
    point = glue::glue("SRID=4326;POINT({longitude} {latitude})")
  )

# db transaction spans entire batch, data only visible in TSC on completed POST
flora_tae10 = flora_tae[1:10,]
wastdr::wastd_POST(flora_tae10, serializer = "occ-taxon-points", api_url = dev)
```

# Data catalogue upload
Snapshots of this workbook (soon: and selected outputs) are uploaded to the data catalogue.

```{r upload_ckan}
d <- ckanr::package_show("threatened-species-and-communities-occurrence-analysis")
# upload this workbook, species EOOs, and comms EOOs.

ckanr::resource_update("56335544-72cf-4ebd-bb2c-bc246ec96080", "occ_etl.html")
```
