---
title: "Occurrence analysis"
author: "Florian Mayer, DBCA"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
# devtools::install_github("ropenscilabs/skimr")
# devtools::install_github("tidyverse/readr")
# install from github: "brianb/mdbtools"
# devtools::install_github("ropensci/ckanr")
library(tidyverse)
library(magrittr)
library(lubridate)
library(skimr)
library(leaflet)
library(Hmisc)
library(ckanr)
library(grDevices)
library(sf)
library(geojsonsf)
library(mapview)
library(ggplot2)

# Configure ckanr to data.dpaw.wa.gov.au
ckanr::ckanr_setup(url = Sys.getenv("CKAN_URL"), key = Sys.getenv("CKAN_API_KEY"))
source("helpers.R")
```

# Context
This workbook analyses occurrence data of threatened species and communities as
extracted from the legacy databases.

Currently, data snapshots from the data catalogue as exported by the ETL workbooks are used. Soon, we will use occurrence data from BioSys.

# Data download
* Now: Occurrence data snapshots from data catalogue
* Soon: Occurrence data from Biosys

```{r}
fauna_occ_rid <- "b73ae0de-f243-4bb3-bbce-25b65bb98e66"
flora_occ_rid <- "b2e7f5a9-8ffb-41c4-bd11-ab090ccb71ef"
comm_occ_rid <- ""
fauna_occ <- ckanr::resource_show(fauna_occ_rid)$url %>% read_csv()
```

# Analysis
## Extent of Occurrence
This section will calculate the Extent of Occurrence (EOO) 
as convex hulls around all occurrences 
of each `name_id` as recorded within the past 10 years.

```{r}
occ_tally <- fauna_occ %>% 
    group_by(name_id, scientific_name) %>% 
    tally() %>% 
    dplyr::arrange(desc(n))
kable(head(occ_tally, n=20))
```
There are `r nrow(occ_tally)` fauna taxa with recorded occurrences.

## Calculating EOO from occurrence point observations
```{r}
#' Return the convex hull around all points with given name_id
#' 
#' @param data A tibble or data.frame with columns longitude, latitude, name_id
#' @param nid A numeric name_id, default: null (use all rows)
#' @param nid_col The column name of the NameID column, default: "name_id"
#' @param lonlat_cols A vector of lon and lat column names, default: c("longitude", "latitude")
#' @param crs The coordinate reference system ID (4283=GDA94)
#' @return An object of class "sfc_POLYGON" "sfc" (package sf)
eoo_polygon <- function(data, 
                        nid=NULL,
                        nid_col="name_id",
                        lonlat_cols=c("longitude", "latitude"),
                        crs=4283,
                        discard=NULL){
    if (is.null(nid)) {
        selected <- data
    } else {
        selected <- dplyr::filter(data, (!! rlang::sym(nid_col)) == nid)
    }
    selected %>% 
        sf::st_as_sf(coords = lonlat_cols, crs = crs, agr = "constant") %>% 
        sf::st_union(.) %>% 
        sf::st_convex_hull(.)
}

eoo_all <- eoo_polygon(fauna_occ)
mapview(eoo_all)

# https://tsc.dbca.wa.gov.au/api/1/taxon/?name_id=24557
# Leipoa ocellata, Malleefowl
eoo_24557 <- eoo_polygon(fauna_occ, nid = 24557)

# see https://r-spatial.github.io/sf/articles/sf5.html
# ggplot() + geom_sf(data = eoo_24557)
mapview(eoo_24557)

# https://tsc.dbca.wa.gov.au/api/1/taxon/?name_id=24734
# Calyptorhynchus latirostris, Carnaby's Black Cockatoo: 14k occurrences
eoo_24734 <- eoo_polygon(fauna_occ, nid = 24734)
mapview(eoo_24734)

# Write sf object to GeoJSON geometry:
# https://cran.r-project.org/web/packages/geojsonsf/vignettes/geojson-sf-conversions.html

# one_taxon_eoo <- tibble(
#     name_id = 24557,
#     eoo = eoo_polygon(fauna_occ, nid = 24557) %>% sfc_geojson()
# )
# wastdr::wastd_POST(one_taxon_eoo, "taxon", verbose=T)
```

# Data upload
## EOO to data catalogue upload
Snapshots of the EOO polygons are uploaded to the data catalogue.
```{r}
d <- ckanr::package_show("wacensus")
# upload this workbook, species EOOs, and comms EOOs.
```

## EOO to TSC upload
This section will upload the EOOs to TSC through its API using `wastdr`.

### Fauna

```{r}
# List unique name_ids without NA and 0
name_ids <- unique(fauna_occ$name_id)
name_ids_nn <- name_ids[!name_ids %in% c(NA, 0)]

# Calculate EOO for each name_id
taxon_eoo <- list()
for (i in seq_len(name_ids_nn)) {
    taxon_eoo[[i]] <- eoo_polygon(fauna_occ, nid = name_ids_nn[[i]]) %>% geojsonsf::sfc_geojson(.)
}

# Assemble tibble of name_id and eoo
taxon_eoos <- tibble(name_id = name_ids_nn, eoo = unlist(taxon_eoo))

# Upload tibble of name_id and eoo to TSC
wastdr::wastd_POST(taxon_eoos, "taxon", verbose = T)
```

### Flora
TODO

### TEC
TODO
