---
title: "Report TFA-TSC: Fauna"
author: "Florian Mayer and Camilla Piggott, DBCA"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE)
source("helpers.R")
```

# Context
[Fauna dataset](https://data.dpaw.wa.gov.au/dataset/threatened-and-priority-fauna-database) 
on the [departmental data catalogue](https://data.dpaw.wa.gov.au/).

[TSC Fauna database](https://tsc.dbca.wa.gov.au/species/?paraphyletic_groups=20&is_terminal_taxon=true) 
on the [TSC catalogue](https://tsc.dbca.wa.gov.au/).

The code for all SCB workbooks is under version control at [github](https://github.com/dbca-wa/scarab-scripts).

```{r tfa_data}
tfa_data<- here::here("data","fauna_cons_listing.csv") %>%
  readr::read_csv(col_types = cols(
      list_code = col_character(),
      criteria_code = col_character(),
      name_id = col_character()))
```

##  Summary statistics

Check unique number of species by both spcies code and scientific name. Check unique conservation criteria and conservation categories

```{r species_summary}
species_unique<-unique(tfa_data$species_code) #1004
scientific_name_unique<-unique(tfa_data$scientific_name) #1004
category_code_unique<-unique(tfa_data$category_code) #27
```

# Taxonomy

## QA: Name mismatches
For a given NameID, the scientific name in TFA should be identical to 
the scientific name in WACensus with the same NameID. 
A name mismatch is an indication for review.

```{r wacensus_names}
wacensus_names <- here::here("data","wace_names.csv") %>%
  readr::read_csv(col_types = cols())

tfa_data_wacensus <- tfa_data %>% dplyr::left_join(wace_names, by="name_id")
```

```{r wacensus_name_mismatch}
name_mismatches <- tfa_data_wacensus %>% 
    dplyr::filter(scientific_name != wace_name) %>% 
    dplyr::select(name_id, species_code, scientific_name, wace_name)
```

### Mismatches containing "subsp"
WACensus names contain "subsp." whereas TFA don't. 
Therefore, scientific name mismatches by only the "subsp" can be ignored.

**QA** Review this list to make sure both WACensus and TFA mean the same name.
If the TFA "scientific_name" means a different taxon than the "wace_name", TFA's
NameID must be changed to the WACensus NameID of the correct taxon.

```{r name_mismatches_subsp}
name_mismatches %>% dplyr::filter(grepl('subsp', wace_name)) %>% DT::datatable()
```

### Remaining mismatches
The remaining mismatches contain minor differences, such as "sp." vs "sp", parentheses and quotation marks.

```{r name_mismatches_typo}
name_mismatches %>% dplyr::filter(!grepl('subsp', wace_name)) %>% DT::datatable()
```

### Remaining mismatches
The remaining mismatches require a review of the scientific name or NameID in TFA.

**QA** If the TFA "scientific_name" is meant to refer to a different taxon than the "wace_name", 
TFA's NameID must be changed to the WACensus NameID of the correct taxon.

```{r name_mismatches_real}
drop_cruft <- . %>% stringr::str_remove_all("[.()'\"]|sp|subsp") %>% stringr::str_replace_all("  ", " ")
true_mismatches <- species_nid %>% 
    dplyr::mutate(
        scientific_name = scientific_name %>% drop_cruft,
        wace_name = wace_name %>% drop_cruft
    ) %>% 
    dplyr::filter(scientific_name != wace_name) %>% 
    dplyr::select(name_id, species_code, short_code, scientific_name, wace_name)

true_mismatches %>% DT::datatable()
```
