---
title: "Report TSC: Fauna"
author: "Milly Piggott and Florian Mayer, DBCA"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE)
source("helpers.R")
```

Make sure to run `wastdr` v 0.1.22 - `remotes::install_github("dbca-wa/wastdr")`.
TODO: this contains fauna and flora. This report is only interested in fauna.

```{r load_data, message=F, warning=F, echo=F}
# TODO load taxa, filter to kingdom==8 (Fauna) and kingdom!=8 (Flora)
# TODO filter tcl to Fauna by right joining it to taxa_fauna
tsc_tcl <- wastdr::wastd_GET("taxon-conservationlisting") %>% 
  wastdr::parse_taxon_conservationlisting()
tsc_tcl %>% head(n=20) %>% knitr::kable()
```

# Context
QA TFA fauna dataset and TSC fauna dataset and make comparisons to check all data has been transfered correctly to TSC.

[TFL Fauna dataset](https://data.dpaw.wa.gov.au/dataset/threatened-and-priority-fauna-database) 
on the [departmental data catalogue](https://data.dpaw.wa.gov.au/).

[TSC Fauna database](https://tsc.dbca.wa.gov.au/species/?paraphyletic_groups=20&is_terminal_taxon=true) 
on the [TSC catalogue](https://tsc.dbca.wa.gov.au/).

The code for all SCB workbooks is under version control at [github](https://github.com/dbca-wa/scarab-scripts).

# TFL fauna conservation listings

```{r tfa_data}
tfa_data<- here::here("data","fauna_cons_listing.csv") %>%
  readr::read_csv(col_types = cols(
      list_code = col_character(),
      criteria_code = col_character(),
      name_id = col_character()))
```

## Summary statistics

Check unique number of species by both spcies code and scientific name. Check unique conservation criteria and conservation categories
```{r species_summary}
species_unique<-unique(tfa_data$species_code) #1004
scientific_name_unique<-unique(tfa_data$scientific_name) #1004
category_code_unique<-unique(tfa_data$category_code) #27
list_code_unique<-unique(tfa_data$list_code) #11
```

There are 4014 unique observations in TFL fauna dataset and 
`r unique(tfa_data$species_code)` unique species. 
There are `r unique(tfa_data$list_code)` unique conservation lists and 
`r unique(tfa_data$category_code)` conservation category codes.

# Taxonomy

## Data ETL QA

### Name mismatches
For a given NameID, the scientific name in TFA should be identical to 
the scientific name in WACensus with the same NameID. 
A name mismatch is an indication for review.

```{r wacensus_names}
wacensus_names <- here::here("data","wace_names.csv") %>%
  readr::read_csv(col_types = cols(
      name_id = col_character()
      )
)
tfa_data_wacensus <- tfa_data %>% dplyr::left_join(wacensus_names, by="name_id")
```

```{r wacensus_name_mismatch}
name_mismatches <- tfa_data_wacensus %>% 
    dplyr::filter(scientific_name != wace_name) %>% 
    dplyr::select(name_id, species_code, scientific_name, wace_name)
```

#### Mismatches containing "subsp"
WACensus names contain "subsp." whereas TFA don't. 
Therefore, scientific name mismatches by only the "subsp" can be ignored.

**QA** Review this list to make sure both WACensus and TFA mean the same name.
If the TFA "scientific_name" means a different taxon than the "wace_name", TFA's
NameID must be changed to the WACensus NameID of the correct taxon.

```{r name_mismatches_subsp}
name_mismatches %>% dplyr::filter(grepl('subsp', wace_name)) %>% DT::datatable()
```

#### Typo mismatches
The remaining mismatches contain minor differences, such as "sp." vs "sp", parentheses and quotation marks.

```{r name_mismatches_typo}
name_mismatches %>% dplyr::filter(!grepl('subsp', wace_name)) %>% DT::datatable()
```

#### True mismatches
The remaining mismatches require a review of the scientific name or NameID in TFA.

**QA** If the TFA "scientific_name" is meant to refer to a different taxon than the "wace_name", 
TFA's NameID must be changed to the WACensus NameID of the correct taxon.

```{r name_mismatches_real}
drop_cruft <- . %>% stringr::str_remove_all("[.()'\"]|sp|subsp") %>% stringr::str_replace_all("  ", " ")
true_mismatches <- tfa_data_wacensus %>% 
    dplyr::mutate(
        scientific_name = scientific_name %>% drop_cruft,
        wace_name = wace_name %>% drop_cruft
    ) %>% 
    dplyr::filter(scientific_name != wace_name) %>% 
    dplyr::select(name_id, species_code, scientific_name, wace_name)

true_mismatches %>% DT::datatable()
```

## Reports
TODO: find reports on Taxonomy. Note that taxa with conservation listings are reported below.

# Conservation Listings

## Data ETL QA

* TSC Taxonomy - get all taxa - filter by kingdom == 8 (Fauna) and anything else (Flora)
* Always compare old TFA against new TSC
* Total numbers cons listings
* Group cons listings by Cons lists, tally
* Anti-join everywhere to first show discrepancies, then (once fixed) prove absence of discrepancies = full equivalence
* Include TSC admin URLs as hyperlinks so that TSC users can view the record live in TSC
* Include `source` (with clear text) and `source_id` = TFA table PKs, taxon `name_id`, taxon shortcode so that TFA custodians can find a record in TFA
* Tell latest conservation listing by `max(effective_from)` date, so we know how fresh the snapshot is

## Reports

### Report to minister for Government Gazette
* Heading: Legislation is BCA
* List taxa with canonical/taxonomic name, group by cons category

### Report the history of one taxon
Use an example taxon (`name_id`) and show:

* Current cons listing in WA (BCA/WAWCA), CWTH (EPBC act)
* Past cons listings in WA, CWTH


## Part content, TODO move up




### Summary statistics

Check unique number of species by both spcies code and scientific name. Check unique conservation criteria and conservation categories
```{r species_summary}
species_unique<-unique(tsc_tcl$species_code) #1004
scientific_name_unique<-unique(tsc_tcl$scientific_name) #1004
category_code_unique<-unique(tsc_tcl$category_code) #27
list_code_unique<-unique(tsc_tcl$list_code) #11
```

There are 4014 unique observations in TFL fauna dataset and 
`r unique(tfa_data$species_code)` unique species. 
There are `r unique(tfa_data$list_code)` unique conservation lists and 
`r unique(tfa_data$category_code)` conservation category codes.


# Occurrences

# Fire History and Actions

# Conservation Threats and Actions
