---
title: "Data ETL from ODK Central: Fauna Spotlighting"
author: "Florian Mayer, DBCA"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
  pdf_document: default
---


# Background
This workbook demonstrates how to access data from ODK Central through the
OData service endpoint.

As an example, we show three forms from Fauna spotlighting surveys:

* Survey start: team, weather conditions at start.
* Spotlighting (repeated for the number of encountered animals): an observer
detects a possum sitting on a tree in the dark by waving a spotlight while
sitting on the roof of a (slowly moving) vehicle.
* Survey end: climate and environment at end of survey.

Data is uploaded to the 
[ODK Central sandbox](https://sandbox.central.opendatakit.org/#/projects/14/).
User and app accounts can be created on request.

This example requires the email address (=ODK Central username) and password 
to be set as R env vars `ODKC_UN` and `ODKC_PW` for the ODK Central server.
A good place for those is `~/.Rprofile`.

```{r setup, message=F}
if (file.exists("~/.Rprofile")) source("~/.Rprofile")
# library(OData)  # didn't work, went with httr
# remotes::install_github("r-lib/rlang")
# remotes::install_github("tidyverse/tidyr")
library(ckanr)
# library(Hmisc)  # MS Access
library(glue)
library(httr)
library(tidyr)
library(dplyr)
library(magrittr)
library(tibble)
library(purrr)
library(leaflet)
library(wastdr)
library(ruODK)
# Our example forms' OData endpoint, plus /Submissions
base_url <- "https://sandbox.central.opendatakit.org/v1/projects/14/forms/"
sss_url <- glue::glue("{base_url}build_Spotlighting-Survey-Start-0-3_1558320795.svc")
sse_url <- glue::glue("{base_url}build_Spotlighting-Survey-End-0-3_1558320208.svc")
spt_url <- glue::glue("{base_url}build_Spotlighting-0-5_1558320001.svc")
```

# Data ETL

## Extract: accessing the data
In this section, form submissions are downloaded from ODK Central and transformed
into a rectangular shape.

```{r}
# unnest_location <- . %>% tidyr::unnest_wider(coordinates) %>%
#   dplyr::rename(longitude=`...1`, latitude=`...2`, altitude=`...3`)

sss_meta <- ruODK::get_metadata(sss_url)
sss_raw <- ruODK::get_submissions(sss_url)
survey_start <- tibble::tibble(value=sss_raw$value) %>% unnest_all

survey_start <- survey_start %>% 
    dplyr::rename(longitude=`...1`, latitude=`...2`, altitude=`...3`) %>% 
  invisible


sse_meta <- ruODK::get_metadata(sse_url)
sse_raw <- ruODK::get_submissions(sse_url)
survey_end <- tibble::tibble(value=sse_raw$value) %>% 
  tidyr::unnest_wider(value) %>% 
  dplyr::rename(uuid=`__id`) %>% 
  tidyr::unnest_wider(`__system`) %>%
  tidyr::unnest_wider(meta) %>%
  tidyr::unnest_wider(location_group) %>%
  tidyr::unnest_wider(location) %>%
  unnest_location %>%
  tidyr::unnest_wider(methods) %>%
  tidyr::unnest_wider(climate) %>%
  invisible

spt_meta <- ruODK::get_metadata(spt_url)
spt_raw <- ruODK::get_submissions(spt_url)
spotlights <- tibble::tibble(value=spt_raw$value) %>% 
  tidyr::unnest_wider(value) %>% 
  dplyr::rename(uuid=`__id`) %>% 
  tidyr::unnest_wider(`__system`) %>%
  tidyr::unnest_wider(meta) %>%
  tidyr::unnest_wider(location_group) %>%
  tidyr::unnest_wider(location) %>%
  unnest_location %>%
  tidyr::unnest_wider(animal_details) %>%
  tidyr::unnest_wider(surroundings) %>%
  invisible
```

For data engineers: The response's content is a list of lists.
One record looks like this:

```{r}
sss_raw$value[[1]]
```

We unnest each nesting level with `tidyr::unnest_wider`, until all terminal
variables are at the same level. This replaces the earlier implementation
using multiple calls to `purrr:map_*` with shorter code - we don't have to 
manually reference each variable combination any longer.


The metadata for one service endpoint looks like this:
```{r}
sss_meta %>% listviewer::jsonedit()
```

We don't do anything with the metadata yet. In future, we could automate the
unnesting based on the dataset schema information in the metadata.

### Data previews
This section demonstrates that we can access, transform, and visualise the data.
This proves that the data is not trapped in ODK Central. 

In production, a section like this can provide a first look at the data (as soon 
as the capture devices have finished their upload).

#### Survey Start
```{r}
survey_start %>% DT::datatable(.)
```

```{r}
leaflet::leaflet(width = 800, height = 600) %>%
  leaflet::addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
  leaflet::addProviderTiles("OpenStreetMap.Mapnik", group = "Place names") %>%
  leaflet::clearBounds() %>% 
  leaflet::addAwesomeMarkers(
    data = survey_start,
    lng = ~longitude, lat = ~latitude,
    icon = leaflet::makeAwesomeIcon(
      text = "S",
      markerColor = ~number_of_spotlights
    ),
    label = ~glue::glue('{survey_start_datetime} {transect_name}'),
    popup = ~glue::glue(
      "<h3>{transect_name}</h3>",
      "Survey start {survey_start_datetime}</br>",
      "<h5>Team</h5>",
      "Observer left {observer_left}</br>",
      "Observer right {observer_right}</br>",
      "Driver {driver}</br>",
      "Recorder {recorder}</br>",
      "<h5>Methodology</h5>",
      "Distance measured by {distance_measure}</br>",
      "Spotlights: {number_of_spotlights}</br>",
      "<h5>Climate</h5>",
      "Air temp {air_temperature} deg C</br>",
      "Wind {wind_speed} m/s</br>",
      "Precipitation {precipitation}</br>",
      "Moon phase {moon_phase}/8</br>",
      "Cloud cover {cloud_cover}/8</br>"
    ),
    clusterOptions = markerClusterOptions()
  ) %>%
  addLayersControl(
    baseGroups = c("Aerial", "Place names"),
    options = layersControlOptions(collapsed = FALSE)
  )
```


#### Survey End
```{r}
survey_end %>% DT::datatable(.)
```

```{r}
leaflet::leaflet(width = 800, height = 600) %>%
  leaflet::addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
  leaflet::addProviderTiles("OpenStreetMap.Mapnik", group = "Place names") %>%
  leaflet::clearBounds() %>% 
  leaflet::addAwesomeMarkers(
    data = survey_end,
    lng = ~longitude, lat = ~latitude,
    icon = leaflet::makeAwesomeIcon(
      text = "E",
      markerColor = "green"
    ),
    label = ~glue::glue('Survey end {survey_end_datetime}'),
    popup = ~glue::glue(
      "<h3>Survey end</h3>",
      "<h5>Methodology</h5>",
      "Avg vehicle speed {average_vehicle_speed_kmh} km/h</br>",
      "Comments: {methodology_comments}</br>",
      "<h5>Climate</h5>",
      "Air temp {air_temperature} deg C</br>",
      "Wind {wind_speed} m/s</br>",
      "Precipitation {precipitation}</br>",
      "Moon phase {moon_phase}/8</br>",
      "Cloud cover {cloud_cover}/8</br>"
    ),
    clusterOptions = markerClusterOptions()
  ) %>%
  addLayersControl(
    baseGroups = c("Aerial", "Place names"),
    options = layersControlOptions(collapsed = FALSE)
  )
```

#### Spotlighted animals
```{r}
spotlights %>% DT::datatable(.)
```

```{r}
leaflet::leaflet(width = 800, height = 600) %>%
  leaflet::addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
  leaflet::addProviderTiles("OpenStreetMap.Mapnik", group = "Place names") %>%
  leaflet::clearBounds() %>% 
  leaflet::addAwesomeMarkers(
    data = spotlights,
    lng = ~longitude, lat = ~latitude,
    icon = leaflet::makeAwesomeIcon(
      text = "A",
      markerColor = ~species
    ),
    label = ~glue::glue('{encounter_start_datetime} {species}'),
    popup = ~glue::glue(
      "<h3>{humanize(species)} ",
      "({humanize(species_id_certainty)}) ({number_of_individuals})</h3>",
      "<h5>Animal</h5>",
      "Dist {distance_to_animal_metres} m, ",
      "bearing {bearing_to_animal_degrees} deg, ",
      "height {animal_height_above_ground_estimate_metres} m</br>",
      "Sitting on {humanize(substrate_under_animal)} / {humanize(tree_species)}</br>",
      "Activity {animal_activities}<br/>",
      "Site type {humanize(site_type)}</br>",
      "<h5>Observer</h5>",
      "Bearing to road {bearing_to_road_degrees} deg<br/>",
      "GPS waypoint {gps_reference_point}<br/>",
      "Odometer {distance_odometer} m<br/>"
    ),
    clusterOptions = markerClusterOptions()
  ) %>%
  addLayersControl(
    baseGroups = c("Aerial", "Place names"),
    options = layersControlOptions(collapsed = FALSE)
  )
```

## Transform: ODK to FaunaFile
Tasks: Translate lookups, rename columns, restructure tables.

FaunaFile has specific internal values for columns with dropdown selections.
To translate the more human-readable ODK values into the legacy FaunaFile values, 
`dplyr::left_join` tibbles of ODK values vs FaunaFile lookups to the data tibbles.

## Load: Upload into FaunaFile
Tasks: Use Hmisc or [RODBC](http://rprogramming.net/connect-to-ms-access-in-r/) to write to an MS Access db. Alteratively, dump the data into CSV, then upload those CSVs into FF
e.g. through an SQL script from inside FF, or import them into the Oracle backend
with a separate SQL script.

In the worst case, we can dump this data into a spreadsheet and use it from there.
In the best case, we can automate the import into FaunaFile.

```{r, eval=F}
dbfile <- "path/to/faunafile.mdb"
con <- Hmisc::mdb.get(dbfile, dateformat = "%m-%d-%Y", as.is = TRUE)
# write transformed data to FaunaFile
```

# Publish
This workbook will be uploaded to the DBCA data catalogue (once productionised).
As long as test data are shown, it is published to RPubs.com.

