---
title: "Data ETL from ODK Central"
author: "Florian Mayer, DBCA"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
  pdf_document: default
---


# Background

Fauna spotlighting surveys have three forms:

* Survey start
* Spotlighting x number of encountered animals
* Survey end

Data is uploaded at the [ODK Central sandbox](https://sandbox.central.opendatakit.org/#/projects/14/).
User and app accounts can be created on request.

Requires R package [Odata](https://cran.r-project.org/package=OData).
Requires R env vars `ODKC_UN` and `ODKC_PW` for the ODK Central server.

```{r setup, message=F}
if (file.exists("~/.Rprofile")) source("~/.Rprofile")
library(OData)
library(ckanr)
library(Hmisc)
library(glue)
library(httr)
library(dplyr)
library(magrittr)
library(tibble)
library(purrr)
library(leaflet)
sss_url <- "https://sandbox.central.opendatakit.org/v1/projects/14/forms/build_Spotlighting-Survey-Start-0-2_1558066890.svc/Submissions"
sse_url <- "https://sandbox.central.opendatakit.org/v1/projects/14/forms/build_Spotlighting-Survey-End-0-2_1558067123.svc/Submissions"
spt_url <- "https://sandbox.central.opendatakit.org/v1/projects/14/forms/build_Spotlighting-0-4_1558064688.svc/Submissions"
```

# Data ETL

## Extract: accessing the data
Read data from ODK Central and beat them into a rectangular shape.

```{r}
# es_sss <- OData::entitySets(glue::glue("{sss_url}?format=json"))  # 501 not implemented
# md_sss <- OData::metadata(glue::glue("{sss_url}"))
auth <- httr::authenticate(Sys.getenv("ODKC_UN"), Sys.getenv("ODKC_PW"))
hdr <- httr::add_headers(Accept = "json")

sss <- httr::GET(sss_url, hdr, auth) %>% httr::content(.)
survey_start <- sss$value %>% {
    tibble::tibble(
    submission_id = purrr::map_chr(., c("__id")),
    submission_datetime = purrr::map_chr(., c("__system", "submissionDate")),
    submitter_name = purrr::map_chr(., c("__system", "submitterName")),
    device_id = purrr::map_chr(., c("device_id")),
    survey_start_datetime = purrr::map_chr(., c("encounter_start_datetime")),
    longitude = purrr::map(., c("location-group", "location", "coordinates")) %>% extract2(1) %>% extract2(1) %>% as.numeric,
    latitude = purrr::map(., c("location-group", "location", "coordinates")) %>% extract2(1) %>% extract2(2) %>% as.numeric,
    gps_reference_point = purrr::map(., c("location-group", "gps_reference_point")),
    transect_name = purrr::map(., c("location-group", "transect_name")),
    observer_left = purrr::map(., c("team", "observer_left")),
    observer_right = purrr::map(., c("team", "observer_right")),
    driver = purrr::map(., c("team", "driver")),
    recorder = purrr::map(., c("team", "recorder")),
    distance_measure = purrr::map_chr(., c("methods", "distance_measure")),
    number_of_spotlights = purrr::map(., c("methods", "number_of_spotlights")) %>% as.integer,
    air_temperature = purrr::map(., c("climate", "air_temperature")) %>% as.numeric,
    wind_speed = purrr::map(., c("climate", "wind_speed")) %>% as.numeric,
    precipitation = purrr::map(., c("climate", "precipitation")),
    moon_phase = purrr::map(., c("climate", "moon_phase")) %>% as.integer,
    cloud_cover = purrr::map(., c("climate", "cloud_cover")) %>% as.integer
  )
}

sse <- httr::GET(sse_url, hdr, auth) %>% httr::content(.)
survey_end <- sse$value %>% {
    tibble::tibble(
    submission_id = purrr::map_chr(., c("__id")),
    submission_datetime = purrr::map_chr(., c("__system", "submissionDate")),
    submitter_name = purrr::map_chr(., c("__system", "submitterName")),
    device_id = purrr::map_chr(., c("device_id")),
    survey_end_datetime = purrr::map_chr(., c("encounter_start_datetime")),
    longitude = purrr::map(sss$value, c("location-group", "location", "coordinates")) %>% extract2(1) %>% extract2(1) %>% as.numeric,
    latitude = purrr::map(sss$value, c("location-group", "location", "coordinates")) %>% extract2(1) %>% extract2(2) %>% as.numeric,
    gps_reference_point = purrr::map(., c("location-group", "gps_reference_point")),
    transect_name = purrr::map(., c("location-group", "transect_name")),
    average_vehicle_speed = purrr::map_chr(., c("methods", "average_vehicle_speed")),
    air_temperature = purrr::map(., c("climate", "air_temperature")) %>% as.numeric,
    wind_speed = purrr::map(., c("climate", "wind_speed")) %>% as.numeric,
    precipitation = purrr::map(., c("climate", "precipitation")),
    moon_phase = purrr::map(., c("climate", "moon_phase")) %>% as.integer,
    cloud_cover = purrr::map(., c("climate", "cloud_cover")) %>% as.integer
  )
}

spt <- httr::GET(spt_url, hdr, auth) %>% httr::content(.)
spotlights <- spt$value %>% {
    tibble::tibble(
    submission_id = purrr::map_chr(., c("__id")),
    submission_datetime = purrr::map_chr(., c("__system", "submissionDate")),
    submitter_name = purrr::map_chr(., c("__system", "submitterName")),
    device_id = purrr::map_chr(., c("device_id")),
    encounter_datetime = purrr::map_chr(., c("encounter_start_datetime")),
    distance_to_animal_metres = purrr::map_chr(., c("animal-details", "distance_to_animal")),
    bearing_to_animal_degrees = purrr::map_chr(., c("animal-details", "bearing_to_animal")),
    bearing_to_road_degrees = purrr::map_chr(., c("animal-details", "bearing_to_road")),
    animal_height_above_ground_estimate_m = purrr::map_chr(., c("animal-details", "animal_height_above_ground_estimate_m")),
    species = purrr::map_chr(., c("animal-details", "species")),
    species_id_certainty = purrr::map_chr(., c("animal-details", "species_id_certainty")),
    number_of_individuals = purrr::map_chr(., c("animal-details", "number_of_individuals")),
    tree_species = purrr::map_chr(., c("surroundings", "tree_species")),
    substrate_under_animal = purrr::map_chr(., c("surroundings", "substrate_under_animal")),
    site_type = purrr::map_chr(., c("surroundings", "site_type")),
    animal_activities = purrr::map_chr(., c("surroundings", "animal_activities")),
    longitude = purrr::map(sss$value, c("location-group", "location", "coordinates")) %>% extract2(1) %>% extract2(1) %>% as.numeric,
    latitude = purrr::map(sss$value, c("location-group", "location", "coordinates")) %>% extract2(1) %>% extract2(2) %>% as.numeric,
    gps_reference_point = purrr::map(., c("location-group", "gps_reference_point")),
    distance_odometer = purrr::map(., c("location-group", "distance_odometer"))
  )
}
```


### Data previews
This section demonstrates that we can access, transform, and visualise the data.
This proves that the data is not trapped in ODK Central. 

#### Survey Start
```{r}
survey_start %>% DT::datatable(.)
```

#### Survey End
```{r}
survey_end %>% DT::datatable(.)
```

#### Spotlighted animals
```{r}
spotlights %>% DT::datatable(.)
```

#### Maps
Note not all data is shown in popups yet.

```{r}
leaflet::leaflet(width = 800, height = 600) %>%
    leaflet::addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
    leaflet::addProviderTiles("OpenStreetMap.Mapnik", group = "Place names") %>%
    leaflet::clearBounds() %>% 
    leaflet::addAwesomeMarkers(
          data = survey_start,
          lng = ~longitude, lat = ~latitude,
          icon = leaflet::makeAwesomeIcon(
            text = "S",
            markerColor = ~number_of_spotlights
          ),
          label = ~glue::glue('{survey_start_datetime} {transect_name}'),
          popup = ~glue::glue(
            "<h3>{transect_name}</h3>",
            "Survey start {survey_start_datetime}</br>",
            "Observer left {observer_left}</br>",
            "Observer right {observer_right}</br>",
            "Driver {driver}</br>",
            "Recorder {recorder}</br>"
          ),
          clusterOptions = markerClusterOptions()
    ) %>% 
    leaflet::addAwesomeMarkers(
          data = survey_end,
          lng = ~longitude, lat = ~latitude,
          icon = leaflet::makeAwesomeIcon(
            text = "E",
            markerColor = "green"
          ),
          label = ~glue::glue('{survey_end_datetime} {transect_name}'),
          popup = ~glue::glue(
            "<h3>{transect_name}</h3>",
            "Survey end {survey_end_datetime}</br>"
            
          ),
          clusterOptions = markerClusterOptions()
    ) %>%
    addLayersControl(
      baseGroups = c("Aerial", "Place names"),
      options = layersControlOptions(collapsed = FALSE)
    )
```

```{r}
leaflet::leaflet(width = 800, height = 600) %>%
    leaflet::addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
    leaflet::addProviderTiles("OpenStreetMap.Mapnik", group = "Place names") %>%
    leaflet::clearBounds() %>% 
    leaflet::addAwesomeMarkers(
          data = spotlights,
          lng = ~longitude, lat = ~latitude,
          icon = leaflet::makeAwesomeIcon(
            text = "A",
            markerColor = ~species
          ),
          label = ~glue::glue('{encounter_datetime} {species}'),
          popup = ~glue::glue(
            "<h3>{species}</h3>",
            "Dist {distance_to_animal_metres}m, bearing {bearing_to_animal_degrees} deg</br>",
            "Sitting on {substrate_under_animal} / {tree_species}</br>",
            "Site type {site_type}</br>"
          ),
          clusterOptions = markerClusterOptions()
    ) %>%
    addLayersControl(
      baseGroups = c("Aerial", "Place names"),
      options = layersControlOptions(collapsed = FALSE)
    )
```

## Transform: ODK to FaunaFile
Tasks: Translate lookups, rename columns, restructure tables.


## Load: Upload into FaunaFile
Tasks: Use Hmisc or [RODBC](http://rprogramming.net/connect-to-ms-access-in-r/) to write to an MS Access db.

In the worst case, we can dump this data into a spreadsheet and use it from there.
In the best case, we can automate the import into FaunaFile.

```{r, eval=F}
dbfile <- "path/to/faunafile.mdb"
con <- Hmisc::mdb.get(dbfile, dateformat = "%m-%d-%Y", as.is = TRUE)
# write transformed data to FaunaFile
```

# Publish
This workbook is uploaded to the DBCA data catalogue.

