---
title: "Report TSC: Fauna"
author: "Milly Piggott and Florian Mayer, DBCA"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 10
    fig_height: 6
    code_folding: hide
    theme: lumen
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE)
source("helpers.R")
```

# Data
The zip archive containing the MS Access database is downloaded from the data 
catalogue, the archive is extracted, and a database connection is opened.

```{r fauna_data_load, message=FALSE}
fauna_data <- dl_mdbzip("66efb68d-8f05-4bfc-af14-5d1a381d0cf2")
```

```{r lookup_tables}
# Lookup tables
shires <- fauna_data$Shires %>%
  dplyr::transmute(
    lga_id = LGACode %>% as.integer(),
    shire_name = LGAName %>% as.character(),
    shire_label = LGAName2 %>% as.character()
  )

tenure <- fauna_data$`Land Tenure` %>%
  dplyr::transmute(
    tenure_id = TenCode %>% as.character(),
    tenure = Tenure %>% as.character()
  )

vegetation_types <- fauna_data$`Vegetation Types` %>%
  dplyr::mutate(
    vegetation_type_id = VegCode %>% as.character(),
    vegetation_type = VegTypeName %>% as.character()
  )

certainty <- fauna_data$Certainty %>%
  dplyr::transmute(
    certainty_id = Code %>% as.integer(),
    certainty = Certainty %>% as.character()
  )

users <- tibble::tibble(
  username = c(
    "AmyM", "AbbyT", "BrianaWingfield",
    "ChristineFreegard", "GeorginaA", "KellieMantle",
    "NickyMarlow", "PeterMawson", "PeterOrell"
  ),
  initials = c(
    "AM", "AT", "BW",
    "CF", "GA", "KM",
    "NM", "PM", "PO"
  ),
  name = c(
    "Amy Mutton", "Abby Thomas", "Briana Wingfield",
    "Christine Freegard", "Georgina Anderson", "Kellie Mantle",
    "Nicky Marlow", "Peter Mawson", "Peter Orell"
  )
)

phylo_class <- fauna_data$`Class List` %>%
  dplyr::transmute(
    phylo_class_id = ClassCode %>% as.character(),
    phylo_class_common_name = ClassName %>% as.character(),
    phylo_class_scientific_name = Class %>% as.character(),
    order = ClassOrder %>% as.character()
  ) %>%
  dplyr::arrange(order)

phylo_group <- fauna_data$`Group List` %>%
  dplyr::transmute(
    phylo_group_id = Group %>% as.integer(),
    phylo_group_name = GroupName %>% as.character(),
    phylo_class_id = Class %>% as.character()
    # order = GroupOrder %>% as.character()
  ) %>%
  dplyr::left_join(phylo_class, by = "phylo_class_id")
# skimr::skim(group)

species_notes <- fauna_data$`Species Notes` %>%
  dplyr::transmute(
    species_code = SpCode %>% as.character(),
    species_notes = Notes %>% as.character()
  )

resolution <- fauna_data$Resolution %>%
  dplyr::transmute(
    resolution_id = ResCode %>% as.integer(),
    resolution_geographic_coords = ResolutionLL %>% as.character(),
    resolution_projected_coords = ResolutionUTM %>% as.character()
  )

resolution_m <- tibble::tibble(
    resolution_id = c(1,2,3,4,5,6),
    accuracy = c(50,100,1000, 10000, 50000, 100000)
)

category <- fauna_data$`Category List` %>%
  dplyr::transmute(
    category_id = Category %>% as.character(),
    phylogenetic_category = CategoryName %>% as.character()
  )
```

```{r fauna_species}
# TODO species notes
species <- fauna_data$`Species List` %>%
  dplyr::transmute(
    species_code = SpCode %>% as.character(),
    old_species_code = OldCode %>% as.character(),
    short_code = ShCode %>% as.character(),
    name_id = NameID %>% as.character(),
    scientific_name = ScName %>% as.character(),
    common_name = ComName %>% as.character(),
    phylo_group_id = PhyloGroup %>% as.integer(),
    family = Family %>% as.character(),
    genus = Genus %>% as.character(),
    species = Species %>% as.character(),
    authority_species = Authority.sp %>% as.character(),
    subspecies = Subspecies %>% as.character(),
    authority_subspecies = Authority.ssp %>% as.character(),
    # file_number_calm = CALMFileNum %>%
    #   as.character() %>%
    #   map(~ as_filenumber(., prefix = "CALM")),
    # file_number_dec = DECFileNum %>%
    #   as.character() %>%
    #   map(~ as_filenumber(., prefix = "DEC")),
    # file_number_dpaw = DPaWFileNum %>%
    #   as.character() %>%
    #   map(~ as_filenumber(., prefix = "DPAW")),
    # file_number_dbca = DBCAFileNum %>%
    #   as.character() %>%
    #   map(~ as_filenumber(., prefix = "DBCA")),
    category_id = Category %>% as.character(),
    origin = Origin %>% as.character(), # N - native, I - introduced
    # region_id_list = Region %>% map(~ as.list(strsplit(., ",")[[1]] %>% as.integer())),
    taxon_id = TaxonId %>% as.logical()
    # translocated = Trans %>% as.character(), # E - extant, T - translocated
    # accept_new_records = Incl %>% as.logical(), # only FALSE in old records
    # species_can_be_trapped = Trap %>% as.logical(),
    # only_sighting_type_surveys = Sight %>% as.logical(),
    # could_have_nest_box = Nbox %>% as.### Datalogical(),
    # could_display_secondary_signs = Sign %>% as.logical(),
    # can_associate_band_number = BANo %>% as.logical(),
    # buffer = Buffer %>% as.character(), # NA 0 1 2 3 - unused
    # estimated_lifespan = EST.LIFE %>% as.integer() # NA  0 10  7 15  6  8
  ) %>%
  dplyr::left_join(phylo_group, by = "phylo_group_id") %>%
  dplyr::left_join(species_notes, by = "species_code") %>%
  dplyr::left_join(category, by = "category_id")
species_skim <- skimr::skim(species)
# print(species_skim)
DT::datatable(species)

# Just the essentials: species code, name_id, names.
species_names <- species %>%
  dplyr::select(species_code, name_id, scientific_name, common_name)
```

```{r tfa_cons_listing}
fauna_cons_listing <- fauna_data$Gazettal %>%
  dplyr::transmute(
    species_code = SpCode %>% as.character(),
    list_code = ListCode %>% as.character(),
    category_code = StatCode %>% as.character(),
    population = Popn %>% as.character(),
    criteria_code = Reasons %>% as.character(),
    authority = AUTHORITY %>% as.character(),
    gazetted_on = DateList %>% parse_date_time(., orders = orders, tz = tz),
    delisted_on = DateDeList %>% parse_date_time(., orders = orders, tz = tz),
    last_reviewed_on = DateReview %>% parse_date_time(., orders = orders, tz = tz),
    reviewed_by = INTERIM %>% as.character(),
    review_outcome = TRANSLOCATION %>% as.character(),
    comments = Comments %>% as.character()
  ) %>%
  dplyr::left_join(species_names, by = "species_code") %>% 
  tibble::rowid_to_column("source_id")
fauna_cons_listing %>% skim()
fauna_cons_listing %>% head(10) %>% DT::datatable()
readr::write_csv(fauna_cons_listing, path = here::here("data","fauna_cons_listing.csv"))
```

```{r compare_tsc_conscrit_listing_to_tfa_conscrit}
chr2int <- . %>% 
      stringr::str_split(",") %>% 
      purrr::map(as.integer)

tfa_conscrit <- here::here("tfa_conscrit_w_tsc.csv") %>%
readr::read_csv(
  col_types = cols(
    list_code = col_character(),
    tsc_criteria = col_character(),
    criteria_code = col_character(),
    assigned_as_in_tsc = col_character(),
    comments_MP = col_character(),
    actions = col_character()
  )
 ) %>%
dplyr::mutate(
  tsc_criteria = tsc_criteria %>% chr2int()
)

fauna_cons_listing_w_tsc <- fauna_cons_listing %>%
  left_join(tfa_conscrit,by=c("list_code","criteria_code")) #4014

# taxon_id_tfa <- fauna_cons_listing_w_tsc %>% 
#   dplyr::transmute(
#     taxon = name_id %>% as.integer()) %>% 
#   distinct() #1003
# 
# tfa_cl <- fauna_cons_listing_w_tsc %>% 
#   dplyr::filter(tsc_criteria != "NULL") %>%
#   dplyr::filter(tsc_criteria != "NA")#805 species with criteria
# 
# taxon_unique_tfa <- tfa_cl %>% 
#   dplyr::transmute(
#     taxon = name_id %>% as.integer()) %>% 
#   distinct() #362 unique species with criteria

```

Make sure to run `wastdr` v 0.1.22 - `remotes::install_github("dbca-wa/wastdr")`.
TODO: this contains fauna and flora. This report is only interested in fauna.

```{r load_data, message=F, warning=F, echo=F}
# TODO load taxa, filter to kingdom==8 (Fauna) and kingdom!=8 (Flora)
# TODO filter tcl to Fauna by right joining it to taxa_fauna
# taxa_raw <- wastdr::wastd_GET(
#   "taxon", 
#   query = list(paraphyletic_groups_icontains=20))
# taxa <- taxa_raw$features %>% tidyr::unnest_wider()

tsc_tcl <- wastdr::wastd_GET("taxon-conservationlisting") %>% 
  wastdr::parse_taxon_conservationlisting() 
tsc_tcl %>% head(n=20) %>% knitr::kable() #3061

# taxon_id_tsc <- tsc_tcl %>% 
#   dplyr::transmute(
#     taxon = taxon %>% as.integer()) %>% 
#   distinct() #870
# 
# tsc_cl <- tsc_tcl %>% 
#   dplyr::filter(criteria != "NA") #626
# 
# taxon_unique_tsc <- tsc_cl %>% 
#   dplyr::transmute(
#     taxon = taxon %>% as.integer()) %>% 
#   distinct() #321
# 
# taxon_missing_tsc <- as.list(setdiff(taxon_unique_tfa$taxon,taxon_unique_tsc$taxon)) #41 missing from tsc
# taxon_missing_tfa <- as.list(setdiff(taxon_unique_tsc$taxon,taxon_unique_tfa$taxon))
```

# Context
QA TFA fauna dataset and TSC fauna dataset and make comparisons to check all data has been transfered correctly to TSC.

[TFL Fauna dataset](https://data.dpaw.wa.gov.au/dataset/threatened-and-priority-fauna-database) 
on the [departmental data catalogue](https://data.dpaw.wa.gov.au/).

[TSC Fauna database](https://tsc.dbca.wa.gov.au/species/?paraphyletic_groups=20&is_terminal_taxon=true) 
on the [TSC catalogue](https://tsc.dbca.wa.gov.au/).

The code for all SCB workbooks is under version control at [github](https://github.com/dbca-wa/scarab-scripts).

# TFL fauna conservation listings

## Summary statistics

Check unique number of species by both spcies code and scientific name. Check unique conservation criteria and conservation categories
```{r species_summary}
species_unique<-unique(tfa_data$species_code) #1004
scientific_name_unique<-unique(tfa_data$scientific_name) #1004
category_code_unique<-unique(tfa_data$category_code) #27
list_code_unique<-unique(tfa_data$list_code) #11
```

There are 4014 unique observations in TFL fauna dataset and 
`r unique(tfa_data$species_code)` unique species. 
There are `r unique(tfa_data$list_code)` unique conservation lists and 
`r unique(tfa_data$category_code)` conservation category codes.

# Taxonomy

## Data ETL QA

### Name mismatches
For a given NameID, the scientific name in TFA should be identical to 
the scientific name in WACensus with the same NameID. 
A name mismatch is an indication for review.

```{r wacensus_names}
wacensus_names <- here::here("data","wace_names.csv") %>%
  readr::read_csv(col_types = cols(
      name_id = col_character()
      )
)
tfa_data_wacensus <- fauna_cons_listing_w_tsc %>% dplyr::left_join(wacensus_names, by="name_id")
```

```{r wacensus_name_mismatch}
name_mismatches <- tfa_data_wacensus %>% 
    dplyr::filter(scientific_name != wace_name) %>% 
    dplyr::select(name_id, species_code, scientific_name, wace_name)
```

#### Mismatches containing "subsp"
WACensus names contain "subsp." whereas TFA don't. 
Therefore, scientific name mismatches by only the "subsp" can be ignored.

**QA** Review this list to make sure both WACensus and TFA mean the same name.
If the TFA "scientific_name" means a different taxon than the "wace_name", TFA's
NameID must be changed to the WACensus NameID of the correct taxon.

```{r name_mismatches_subsp}
name_mismatches %>% dplyr::filter(grepl('subsp', wace_name)) %>% DT::datatable()
```

#### Typo mismatches
The remaining mismatches contain minor differences, such as "sp." vs "sp", parentheses and quotation marks.

```{r name_mismatches_typo}
name_mismatches %>% dplyr::filter(!grepl('subsp', wace_name)) %>% DT::datatable()
```

#### True mismatches
The remaining mismatches require a review of the scientific name or NameID in TFA.

**QA** If the TFA "scientific_name" is meant to refer to a different taxon than the "wace_name", 
TFA's NameID must be changed to the WACensus NameID of the correct taxon.

```{r name_mismatches_real}
drop_cruft <- . %>% stringr::str_remove_all("[.()'\"]|sp|subsp") %>% stringr::str_replace_all("  ", " ")
true_mismatches <- tfa_data_wacensus %>% 
    dplyr::mutate(
        scientific_name = scientific_name %>% drop_cruft,
        wace_name = wace_name %>% drop_cruft
    ) %>% 
    dplyr::filter(scientific_name != wace_name) %>% 
    dplyr::select(name_id, species_code, scientific_name, wace_name)

true_mismatches %>% DT::datatable()
```

## Reports
TODO: find reports on Taxonomy. Note that taxa with conservation listings are reported below.

# Conservation Listings

## Data ETL QA

* TSC Taxonomy - get all taxa - filter by kingdom == 8 (Fauna) and anything else (Flora)
* Always compare old TFA against new TSC
* Total numbers cons listings
* Group cons listings by Cons lists, tally
* Anti-join everywhere to first show discrepancies, then (once fixed) prove absence of discrepancies = full equivalence
* Include TSC admin URLs as hyperlinks so that TSC users can view the record live in TSC
* Include `source` (with clear text) and `source_id` = TFA table PKs, taxon `name_id`, taxon shortcode so that TFA custodians can find a record in TFA
* Tell latest conservation listing by `max(effective_from)` date, so we know how fresh the snapshot is

## Reports

### Report to minister for Government Gazette
* Heading: Legislation is BCA
* List taxa with canonical/taxonomic name, group by cons category

### Report the history of one taxon
Use an example taxon (`name_id`) and show:

* Current cons listing in WA (BCA/WAWCA), CWTH (EPBC act)
* Past cons listings in WA, CWTH


## Part content, TODO move up




### Summary statistics

Check unique number of species by both spcies code and scientific name. Check unique conservation criteria and conservation categories
```{r species_summary}
species_unique<-unique(tsc_tcl$species_code) #1004
scientific_name_unique<-unique(tsc_tcl$scientific_name) #1004
category_code_unique<-unique(tsc_tcl$category_code) #27
list_code_unique<-unique(tsc_tcl$list_code) #11
```

There are 4014 unique observations in TFL fauna dataset and 
`r unique(tfa_data$species_code)` unique species. 
There are `r unique(tfa_data$list_code)` unique conservation lists and 
`r unique(tfa_data$category_code)` conservation category codes.


# Occurrences

# Fire History and Actions

# Conservation Threats and Actions

# Management Plans

